### 2017-02-17 doc/adaAbstract.Rmd
ADA Abstract 2017

---
title: "ADA Abstract"
author: "Windy Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(dplyr)

ds <- readRDS(file='../data/ds.Rds') %>% 
  dplyr::mutate(logUdbpRatio = log(udbpCrRatio)) %>% 
  dplyr::select(SID, VN, fVN, Age, ACR, eGFR, UDBP,
                acr_status, eGFR_status, UDBP_status, 
                udbpCrRatio, logUdbpRatio, 
                DM, dm_status)
  # dplyr::group_by(SID) %>% 
  # dplyr::filter(n() == 3) %>% 
  # dplyr::ungroup()

ds1 <- ds %>% 
  dplyr::filter(fVN == "Baseline")

source('../R/functions.r')
source("../R/gee.R")
source("../R/graphs.R")

```

Subject Characterization
========================

Subject characteristics across visit numbers
```{r subject characteristics}
table_subChar(ds1)

# check n
table(ds1$acr_status)

# 
# # check n of multiple variables
ds %>%
  dplyr::tbl_df() %>%
  dplyr::filter(VN == 3) %>%
  dplyr::select(VN, fVN, SID, Age, ACR, eGFR, UDBP) %>%
  tidyr::gather(Measure, Value, -SID) %>%
  na.omit() %>%
  dplyr::group_by(Measure) %>%
  dplyr::summarise(n = n())

```

Statististical tests
```{r stats}

## make sure that it is okay to do chi-square with longitudinal
## make sure the error term for anova is correct

summary(aov(Age ~ fVN + Error(SID/fVN), data=ds))

chisq.test(table(ds$Ethnicity, ds$fVN), correct = TRUE)

chisq.test(table(ds$Sex, ds$fVN), correct = TRUE)

summary(aov(BMI ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Waist ~ fVN + Error(SID/fVN), data=ds))

summary(aov(eGFR ~ fVN + Error(SID/fVN), data=ds))

summary(aov(ACR ~ fVN + Error(SID/fVN), data=ds))

summary(aov(UrineCreatinine ~ fVN + Error(SID/fVN), data=ds))

summary(aov(UrineMicroalbumin ~ fVN + Error(SID/fVN), data=ds))

summary(aov(UDBP ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Creatinine ~ fVN + Error(SID/fVN), data=ds))

summary(aov(VitaminD ~ fVN + Error(SID/fVN), data=ds))

summary(aov(PTH ~ fVN + Error(SID/fVN), data=ds))

summary(aov(ALT ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Systolic ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Diastolic ~ fVN + Error(SID/fVN), data=ds))

summary(aov(MeanArtPressure ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Glucose0 ~ fVN + Error(SID/fVN), data=ds))

summary(aov(Glucose120 ~ fVN + Error(SID/fVN), data=ds))

chisq.test(table(ds$dm_status, ds$fVN), correct = TRUE)

# TukeyHSD(anova, which="fVN")
```

Part 1 - Cross-Sectional
========================

ACR
---
```{r acr boxplot}

# Box plot of uVDBP in different albuminuria categories

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(acr_status, udbpCrRatio) %>%
  na.omit() %>%
  box_plot("acr_status", "log(udbpCrRatio)", 
            "Albuminuria",
            "log uVDBP:Creatinine")

# Boxplot of ACR in uVDBP categories
ds1 %>% 
  select(UDBP_status, ACR) %>% 
  na.omit() %>% 
  box_plot("UDBP_status", "log(ACR)",
           "Urinary VDBP status",
           "log urine albumin-to-creatinine ratio (mg/mmol)")

# Values
table_value(ds1, by = "acr_status")

# ANOVA

anova <- aov(formula = log(udbpCrRatio)~acr_status, data = ds1)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r acr scatterplot}

# Scatterplot of ACR and uVDBP ----------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(ACR, udbpCrRatio) %>% 
  na.omit() %>% 
  scatter_plot("log(ACR)", "log(udbpCrRatio)",
               "log Albumin:Creatinine Ratio",
               "log UDBP:Creatinine") +
  geom_smooth(method = lm, colour = "grey50")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ ACR + udbpCrRatio, data = ., method = "spearman")
```

eGFR
----
```{r eGFR boxplot}

# Boxplot of uVDBP concentrations across eGFR categories --------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(eGFR_status, udbpCrRatio) %>% 
  na.omit() %>% 
  box_plot2("eGFR_status", "log(udbpCrRatio)", 
            "Estimated GFR (ml/min/1.73m^2)",
            "log uVDBP:Creatinine")

# Values
table_value(ds1, by = "eGFR_status")

# ANOVA

anova <- aov(formula = log(udbpCrRatio)~eGFR_status, data = ds1)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r eGFR scatterplot}

# Scatterplot of eGFR and uVDBP ----------------------------------

ds %>% 
  dplyr::filter(!(eGFR_status == "Moderate")) %>% 
  dplyr::filter(eGFR < 150) %>% 
  dplyr::select(eGFR, udbpCrRatio) %>% 
  na.omit() %>% 
  scatter_plot("eGFR", "log(udbpCrRatio)",
               "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
               "log UDBP:Creatinine") +
  ggplot2::geom_smooth(method = lm, colour = "grey50")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(!(acr_status == "Macroalbuminuria")) %>% 
  cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")
```

```{r hyperfiltration}
ds %>% 
  dplyr::filter(eGFR < 45) %>% 
  dplyr::select(SID, VN, Age, eGFR, dm_status, acr_status)

ds %>% 
  scatter_plot("Age", "eGFR",
               "Age", "eGFR")
```

Part 2 - GEE
============

```{r correlation}
dsGEE <- ds %>% 
  dplyr::mutate(udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
                ageBase = ifelse(fVN == "Baseline", Age, NA),
                bmiBase = ifelse(fVN == "Baseline", BMI, NA),
                waistBase = ifelse(fVN == "Baseline", Waist, NA), 
                Sex = ifelse(Sex == "Male", 1, 0),
                Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
                Ethnicity = ifelse(Ethnicity == "European", 0, 1),
                fMedsBP = ifelse(fMedsBP == "Yes", 1, 0)
                # vitdBase = ifelse(fVN == "6Year", NA, VitaminD)
                ) %>% 
  dplyr::arrange(SID, fVN) %>% 
  dplyr::group_by(SID) %>% 
  tidyr::fill(udbpBase, ageBase, bmiBase, waistBase) %>% 
  dplyr::ungroup() %>%
  dplyr::arrange(SID, VN) %>% 
  dplyr::select(udbpBase, ACR, eGFR, VN, ageBase, Sex, Ethnicity, BMI, 
                Waist, DM, fMedsBP)

round(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"), 2)

# Heatmap

corrplot::corrplot(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"),
                   method = "color")

```

```{r mason}

gee <- ds %>%
  dplyr::mutate(
  udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
  ageBase = ifelse(fVN == "Baseline", Age, NA),
  fDM = relevel(as.factor(DM), "1")
  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
  ) %>%
  dplyr::filter(!(fVN == "Baseline" &
                    acr_status == "Macroalbuminuria")) %>%
  dplyr::filter(!(fVN == "Baseline" & eGFR_status == "Moderate")) %>%
  dplyr::arrange(SID, fVN) %>%
  dplyr::group_by(SID) %>%
  tidyr::fill(udbpBase, ageBase) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(UDBP = UDBP/1000) %>%
  dplyr::arrange(SID, VN) %>%
  mason_gee(
  yvars = c("ACR", "eGFR"),
  xvars = "UDBP",
  covars = c("VN", "ageBase", "fDM")
  )
```

```{r}
gee3 <- readRDS(file = "../data/ds.Rds") %>%
  dplyr::mutate(
  udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
  ageBase = ifelse(fVN == "Baseline", Age, NA),
  VitaminD3 = ifelse(fVN == "6Year", NA, VitaminD)
  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
  ) %>%
  dplyr::arrange(SID, fVN) %>%
  dplyr::group_by(SID) %>%
  tidyr::fill(udbpBase, ageBase) %>%
  dplyr::ungroup() %>%
  # dplyr::group_by(VN) %>%
  # dplyr::mutate(udbpBase = as.numeric(scale(udbpBase))) %>%
  dplyr::mutate(UDBP = UDBP/1000) %>%
  # dplyr::ungroup() %>% 
  dplyr::arrange(SID, VN) %>%
  mason_gee(
  yvars = c("VitaminD3"),
  xvars = "UDBP",
  covars = c("VN", "ageBase", "DM")
  )
```

Graph
```{r gee plot}

gee %>%
  dplyr::mutate(p.value = ifelse(p.value > 0.05, 1, 0.04)) %>%
  seer::view_main_effect(
  graph.options = 'dot.size',
  groups = '~Yterms',
  legend.title = 'p-values',
  xlab = 'Percent difference with 95% CI in the outcomes\nfor each unit increase in baseline uVDBP',
  ylab = 'Kidney Measures'
  ) +
  graph_theme(ticks = FALSE, legend.pos = 'right') +
  ggplot2::theme(legend.margin = grid::unit(0, 'cm'))

```

### 2017-02-17 R/load_data.R
Changed project.name to urinaryDBP

@@ -2,7 +2,7 @@
 #'
 #' @param update Whether you want to update the dataset using the fetch_data
 #'   function.
-#' @param urinaryDBP Name of your project. Set the default to keep it always
+#' @param project.name Name of your project. Set the default to keep it always
 #'   working.
 #'
 #' @return Outputs the datasets in the data/ folder.
@@ -12,10 +12,10 @@
 #' load_data()
 #' load_data(update = TRUE)
 #'
-load_data <- function(update = FALSE, urinaryDBP = NULL) {
-    if (is.null(urinaryDBP))
+load_data <- function(update = FALSE, project.name = "urinaryDBP") {
+    if (is.null(project.name))
         stop('Please change the project.name default to the name of the project.')
-    if (!dir.exists(system.file('data', package = urinaryDBP)))
+    if (!dir.exists(system.file('data', package = project.name)))
         fetch_data()
     if (update)
         fetch_data()

### 2017-02-17 doc/adaAbstract.Rmd
Replaced setup with functions and added sub char table

@@
 ---
-title: "ADA Abstract"
-author: "Windy Wang"
-output: html_document
+title: "**Urinary Vitamin D Binding Protein as a Potential Biomarker for Nephropathy in Patients at risk for Type 2 Diabetes**: The Prospective Metabolism and Islet Cell Evaluation (PROMISE) cohort study."
+author: "Wang WZN, Cole DEC, Johnston LW, Wong BYL, Fu L, Retnakaran R, Harris SB, Zinman B, Hanley AJ"
+output: word_document:
+          fig_caption: yes
+          reference_docx: wordStyle.docx
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
+knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
+library(dplyr)
+devtools::load_all()
+load_data(update = TRUE)
+set_options()
+```
 
-library(dplyr)
+```{r dsSetup}
+ds <- project_data
 
-ds <- readRDS(file='../data/ds.Rds') %>% 
-  dplyr::mutate(logUdbpRatio = log(udbpCrRatio)) %>% 
-  dplyr::select(SID, VN, fVN, Age, ACR, eGFR, UDBP,
-                acr_status, eGFR_status, UDBP_status, 
-                udbpCrRatio, logUdbpRatio, 
-                DM, dm_status)
-  # dplyr::group_by(SID) %>% 
-  # dplyr::filter(n() == 3) %>% 
-  # dplyr::ungroup()
-
-ds1 <- ds %>% 
+dsBase <- ds %>% 
   dplyr::filter(fVN == "Baseline")
 
-source('../R/functions.r')
-source("../R/gee.R")
-source("../R/graphs.R")
+dsComplete <- ds %>% 
+  dplyr::group_by(SID) %>%
+  dplyr::filter(n() == 3) %>%
+  dplyr::ungroup()
+```
 
+```{r captionSetup}
+## Include captions below using `captioner` package
+figNums <- captioner::captioner(prefix = 'Figure')
+supFigNums <- captioner::captioner(prefix = 'Supplementary Figure')
+tabNums <- captioner::captioner(prefix = 'Table')
+supTabNums <- captioner::captioner(prefix = 'Supplementary Table')
 ```
@@
 ```{r subject characteristics}
-table_subChar(ds1)
 
-# check n
-table(ds1$acr_status)
-
-# 
-# # check n of multiple variables
-ds %>%
-  dplyr::tbl_df() %>%
-  dplyr::filter(VN == 3) %>%
-  dplyr::select(VN, fVN, SID, Age, ACR, eGFR, UDBP) %>%
-  tidyr::gather(Measure, Value, -SID) %>%
-  na.omit() %>%
-  dplyr::group_by(Measure) %>%
-  dplyr::summarise(n = n())
-
-```
-
-Statististical tests
-```{r stats}
-
-## make sure that it is okay to do chi-square with longitudinal
-## make sure the error term for anova is correct
-
-summary(aov(Age ~ fVN + Error(SID/fVN), data=ds))
-
-chisq.test(table(ds$Ethnicity, ds$fVN), correct = TRUE)
-
-chisq.test(table(ds$Sex, ds$fVN), correct = TRUE)
-
-summary(aov(BMI ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Waist ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(eGFR ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(ACR ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(UrineCreatinine ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(UrineMicroalbumin ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(UDBP ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Creatinine ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(VitaminD ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(PTH ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(ALT ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Systolic ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Diastolic ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(MeanArtPressure ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Glucose0 ~ fVN + Error(SID/fVN), data=ds))
-
-summary(aov(Glucose120 ~ fVN + Error(SID/fVN), data=ds))
-
-chisq.test(table(ds$dm_status, ds$fVN), correct = TRUE)
-
-# TukeyHSD(anova, which="fVN")
+subCharTable <- tableone::CreateTableOne(
+  vars = c("Age",
+           "Sex",
+           "Ethnicity",
+           "BMI",
+           "Waist",
+           "eGFR",
+           "ACR",
+           "UrineCreatinine",
+           "UrineMicroalbumin",
+           "UrinaryCalcium",
+           "UDBP",
+           "Creatinine",
+           "VitaminD",
+           "PTH",
+           "ALT",
+           "Systolic",
+           "Diastolic",
+           "MeanArtPressure",
+           "Glucose0",
+           "Glucose120",
+           "dmStatus",
+           "SmokeCigs"),
+  strata = c("fVN"),
+  data = ds,
+  factorVars = c("Sex", "SmokeCigs")
+) %>% 
+  print(nonnormal = c("Glucose0",
+                      "Glucose120",
+                      "ALT",
+                      "UDBP",
+                      "ACR")) %>% 
+  pander::pander()
 ```

### 2017-02-17 R/fetch_data.R
Fixed syntax error on ACR Status

@@ -52,7 +52,7 @@ fetch_data <- function() {
                "Microalbuminuria")
       ),
       acrStatus = factor(
-        acr_status,
+        acrStatus,
         levels = c("Normoalbuminuria", "Microalbuminuria", "Macroalbuminuria"),
         ordered = TRUE
       ),

### 2017-02-17 R/rename.R
Rename variables in data set to improve understanding and include units

@@ -1,0 +1,31 @@
+
+# Renaming ----------------------------------------------------------------
+
+#' Renames variables in data set to improve understanding and include units
+#'
+#' @param x
+
+#' @export
+#'
+#' @examples
+rename_table_rows <- function(x) {
+  gsub('Age', 'Age (years)', x) %>%
+    gsub('Waist', 'Waist Circumference (cm)', .) %>%
+    gsub('eGFR', 'Estimated GFR (ml/min/1.73m^2)', .) %>%
+    gsub('ACR', 'Urinary Albumin-to-Creatinine Ratio (mg/mmol)', .) %>%
+    gsub('MicroalbCreatRatio', 'Microalbumin:Creatinine', .) %>%
+    gsub('UrineCreatinine', 'Urinary Creatinine (mmol/L)', .) %>%
+    gsub('UrineMicroalbumin', 'Urinary Microalbumin (mg/L)', .) %>%
+    gsub('VitaminD', 'Serum 25(OH)D (nmol/L)', .) %>%
+    gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
+    gsub('Diastolic', 'Diastolic Blood Pressure (mmHg)', .) %>%
+    gsub('MeanArtPressure', 'Mean Arterial Pressure (mmHg)', .) %>%
+    gsub('Systolic', 'Systolic Blood Pressure (mmHg)', .) %>%
+    gsub('PTH', 'Parathyroid Hormone (pmol/L)', .) %>%
+    gsub('ALT', 'Serum ALT (U/L)', .) %>%
+    gsub('Glucose0', 'Fasting', .) %>%
+    gsub('Glucose120', '2h OGTT', .) %>%
+    gsub('dm_status', 'Diabetic Status', .) %>%
+    gsub('DM', 'Diabetes', .) %>%
+    gsub('NGT', 'Normal', .)
+}

### 2017-02-17 R/plots.R
Fixed scatterplot colours

@@ -83,8 +83,6 @@ scatter_plot = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet =
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "grey50"), #angle = 45
                    axis.title = ggplot2::element_text(size = 10)) +
-    ggplot2::scale_fill_brewer(palette = "Dark2") +
-    ggplot2::scale_color_brewer(palette = "Dark2") +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
   if(line == TRUE & facet == FALSE) {

### 2017-02-17 doc/adaAbstract.Rmd
ACR boxplots

@@
 
-ACR
----
-```{r acr boxplot}
+## ACR
+
+```{r acr}
 
 # Box plot of uVDBP in different albuminuria categories
-
 ds %>% 
   filter(fVN == "Baseline") %>% 
-  select(acr_status, udbpCrRatio) %>%
+  select(acrStatus, udbpCrRatio) %>%
   na.omit() %>%
-  box_plot("acr_status", "log(udbpCrRatio)", 
+  box_plot("acrStatus", "log(udbpCrRatio)", 
             "Albuminuria",
@@
 # Boxplot of ACR in uVDBP categories
-ds1 %>% 
-  select(UDBP_status, ACR) %>% 
+dsBase %>% 
+  select(UDBPStatus, ACR) %>% 
   na.omit() %>% 
-  box_plot("UDBP_status", "log(ACR)",
+  box_plot("UDBPStatus", "log(ACR)",
            "Urinary VDBP status",
-           "log urine albumin-to-creatinine ratio (mg/mmol)")
-
-# Values
-table_value(ds1, by = "acr_status")
+           "log Urine ACR (mg/mmol)")
 
 # ANOVA
-
 anova <- aov(formula = log(udbpCrRatio)~acr_status, data = ds1)
@@
                "log Albumin:Creatinine Ratio",
-               "log UDBP:Creatinine") +
-  geom_smooth(method = lm, colour = "grey50")
+               "log UDBP:Creatinine",
+               line = TRUE)
 
@@
 
-eGFR
-----
+## eGFR
+
 ```{r eGFR boxplot}
@@
             "log uVDBP:Creatinine")
-
-# Values
-table_value(ds1, by = "eGFR_status")
 

### 2017-03-01 doc/seminar2017.Rmd
Graphs for my introductory seminar slides

--- 2017-02-17 doc/adaAbstract.Rmd
@@
 ---
-title: "**Urinary Vitamin D Binding Protein as a Potential Biomarker for Nephropathy in Patients at risk for Type 2 Diabetes**: The Prospective Metabolism and Islet Cell Evaluation (PROMISE) cohort study."
-author: "Wang WZN, Cole DEC, Johnston LW, Wong BYL, Fu L, Retnakaran R, Harris SB, Zinman B, Hanley AJ"
-output: word_document:
-          fig_caption: yes
-          reference_docx: wordStyle.docx
+title: "Intro Seminar"
+author: "Windy Wang"
+date: "March 1, 2017"
+output: html_document
 ---
@@
 ds <- project_data
-
-dsBase <- ds %>% 
-  dplyr::filter(fVN == "Baseline")
-
-dsComplete <- ds %>% 
-  dplyr::group_by(SID) %>%
-  dplyr::filter(n() == 3) %>%
-  dplyr::ungroup()
-```
-
-```{r captionSetup}
-## Include captions below using `captioner` package
-figNums <- captioner::captioner(prefix = 'Figure')
-supFigNums <- captioner::captioner(prefix = 'Supplementary Figure')
-tabNums <- captioner::captioner(prefix = 'Table')
-supTabNums <- captioner::captioner(prefix = 'Supplementary Table')
+# 
+# dsBase <- ds %>% 
+#   dplyr::filter(fVN == "Baseline")
+# 
+# dsComplete <- ds %>% 
+#   dplyr::group_by(SID) %>%
+#   dplyr::filter(n() == 3) %>%
+#   dplyr::ungroup()
 ```
@@
            "SmokeCigs"),
-  strata = c("fVN"),
+  strata = c("UDBPStatus"),
   data = ds,
@@
                       "ACR")) %>% 
-  pander::pander()
+  knitr::kable()
 ```

### 2017-03-01 R/plots.R
Boxplot where colour theme matches health template slides

@@ -56,6 +56,52 @@ box_plot <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
 }
 
 
+#' Boxplots where the colour theme matches health template slides
+#'
+#' @param data
+#' @param xvar
+#' @param yvar
+#' @param xlab
+#' @param ylab
+#' @param facet
+#'
+#' @return Plot
+#'
+#' @examples
+box_plot_slides <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
+  myboxplot <- ggplot2::ggplot(data, ggplot2::aes_string(x = xvar, y = yvar)) +
+    ggplot2::geom_boxplot(ggplot2::aes_string(colour = xvar, fill = xvar),
+                          outlier.shape = NA) +
+    ggplot2::stat_summary(geom = "crossbar", width = 0.65, fatten = 0, color = "white",
+                          fun.data = function(x){
+                            return(c(y = median(x), ymin = median(x), ymax = median(x)))
+                          }) +
+    ggplot2::geom_jitter(ggplot2::aes_string(),
+                         position = ggplot2::position_jitter(width = 0.15, height = 0),
+                         alpha = 0.3,
+                         size = 1,
+                         colour = "grey30") +
+    ggplot2::theme_minimal() +
+    ggplot2::scale_color_manual(values = c("#0db7c4", "#f24745", "#a9d039")) +
+    ggplot2::scale_fill_manual(values = c("#0db7c4", "#f24745", "#a9d039")) +
+    ggplot2::theme(legend.position = "none",
+                   panel.grid.major.x = ggplot2::element_blank(),
+                   axis.line.y = ggplot2::element_blank(),
+                   axis.text.y = ggplot2::element_text(colour = "grey"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
+                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
+                   axis.title = ggplot2::element_text(size = 10)) +
+    ggplot2::xlab(xlab) +
+    ggplot2::ylab(ylab)
+
+  if(facet == TRUE) {
+    myboxplot +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else {
+    myboxplot
+  }
+}
+
 
 # Scatterplot -------------------------------------------------------------
 

### 2017-03-01 R/fetch_data.R
Reversed order of eGFRStatus

@@ -61,8 +61,10 @@ fetch_data <- function() {
                                     1, Age, isAfrican),
       eGFRStatus = cut(eGFR,
                        breaks = c(-Inf, 60, 90, Inf),
-                       labels = c("Moderate", "Mild", "Normal"),
-                       ordered_result = TRUE),
+                       labels = c("Moderate", "Mild", "Normal")),
+      eGFRStatus = factor(eGFRStatus,
+                          levels = c("Normal", "Mild", "Moderate"),
+                          ordered = TRUE),
       UDBPStatus = cut(UDBP,
                        breaks = c(0, 1.23, 60, Inf),
                        labels = c("Trace", "Normal", "High"),

### 2017-03-01 R/plots.R
Increased font size for readability in slides

@@ -87,10 +87,10 @@ box_plot_slides <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   axis.text.y = ggplot2::element_text(colour = "grey"),
+                   axis.text.y = ggplot2::element_text(colour = "grey", size = 12),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
-                   axis.title = ggplot2::element_text(size = 10)) +
+                   axis.text.x = ggplot2::element_text(colour = "grey30", size = 12), #angle = 45
+                   axis.title = ggplot2::element_text(size = 14)) +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 

### 2017-03-01 doc/seminar2017.Rmd
Changed box plots to slides theme

@@
   na.omit() %>%
-  box_plot("acrStatus", "log(udbpCrRatio)", 
+  box_plot_slides("acrStatus", "log(udbpCrRatio)", 
             "Albuminuria",
             "log uVDBP:Creatinine")
-
-# Boxplot of ACR in uVDBP categories
-dsBase %>% 
-  select(UDBPStatus, ACR) %>% 
-  na.omit() %>% 
-  box_plot("UDBPStatus", "log(ACR)",
-           "Urinary VDBP status",
-           "log Urine ACR (mg/mmol)")
 
@@
   filter(fVN == "Baseline") %>% 
-  select(eGFR_status, udbpCrRatio) %>% 
+  select(eGFRStatus, udbpCrRatio) %>% 
   na.omit() %>% 
-  box_plot2("eGFR_status", "log(udbpCrRatio)", 
+  box_plot_slides("eGFRStatus", "log(udbpCrRatio)", 
             "Estimated GFR (ml/min/1.73m^2)",
@@
                "Age", "eGFR")
+```
+
+## Glycemic Status
+
+```{r DM boxplot}
+
+# Boxplot of uVDBP concentrations across DM categories
+
+ds %>% 
+  filter(fVN == "Baseline") %>% 
+  select(dmStatus, udbpCrRatio) %>% 
+  na.omit() %>% 
+  box_plot_slides("dmStatus", "log(udbpCrRatio)", 
+            "Glycemic Status",
+            "log uVDBP:Creatinine")
+
+# ANOVA
+
+anova <- aov(formula = log(udbpCrRatio)~eGFR_status, data = ds1)
+summary(anova)
+TukeyHSD(anova)
+rm(anova)
 ```

### 2017-03-03 doc/seminar2017.Rmd
Included birth control in table 1

@@
            "dmStatus",
-           "SmokeCigs"),
-  strata = c("UDBPStatus"),
+           "SmokeCigs",
+           "BirthControl"),
+  strata = c("fVN"),
   data = ds,
-  factorVars = c("Sex", "SmokeCigs")
+  factorVars = c("Sex", "SmokeCigs", "BirthControl")
 ) %>% 

### 2017-03-12 doc/seminar2017.Rmd
Fixed acr variable syntax error
Added baseline 25(OH)D scatterplot

@@
 ds %>% 
-  filter(!(acr_status == "Macroalbuminuria")) %>% 
+  filter(!(acrStatus == "Macroalbuminuria")) %>% 
   cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")
@@
 rm(anova)
+```
+
+## Vitamin D
+
+```{r VitD scatterplot}
+
+ds %>% 
+  dplyr::filter(fVN == "Baseline") %>% 
+  dplyr::select(VitaminD, udbpCrRatio) %>% 
+  scatter_plot("log(udbpCrRatio)", "VitaminD",
+               "log UDBP:Creatinine",
+               "Serum 25(OH)D",
+               line = TRUE)
+
+# Spearman Correlation ------------------------------------------
+
+ds %>% 
+  dplyr::filter(fVN == "Baseline") %>% 
+  cor.test(formula = ~ udbpCrRatio + VitaminD, data = ., method = "spearman")
 ```

### 2017-03-12 doc/seminar2017.Rmd
eGFR is not lower in people with DM

@@
   dplyr::filter(eGFR < 45) %>% 
-  dplyr::select(SID, VN, Age, eGFR, dm_status, acr_status)
+  dplyr::select(SID, VN, Age, eGFR, dmStatus, acrStatus)
 
@@
                "Age", "eGFR")
+
+ds %>% 
+  dplyr::select(dmStatus, eGFR, fVN) %>% 
+  na.omit() %>% 
+  box_plot("dmStatus", "eGFR",
+               "Diabetic Status", "eGFR",
+           facet = TRUE)
 ```

### 2017-03-12 doc/adaAbstract.Rmd
Changed syntax of variables

@@
   dplyr::filter(!(fVN == "Baseline" &
-                    acr_status == "Macroalbuminuria")) %>%
-  dplyr::filter(!(fVN == "Baseline" & eGFR_status == "Moderate")) %>%
+                    acrStatus == "Macroalbuminuria")) %>%
+  dplyr::filter(!(fVN == "Baseline" & eGFRStatus == "Moderate")) %>%
   dplyr::arrange(SID, fVN) %>%

### 2017-03-12 doc/seminar2017.Rmd
Fixed gee coding

@@
            "UrineMicroalbumin",
-           "UrinaryCalcium",
            "UDBP",
@@
            "dmStatus",
-           "SmokeCigs",
-           "BirthControl"),
+           "SmokeCigs"),
   strata = c("fVN"),
   data = ds,
-  factorVars = c("Sex", "SmokeCigs", "BirthControl")
+  factorVars = c("Sex", "SmokeCigs")
 ) %>% 
@@
                       "UDBP",
-                      "ACR")) %>% 
-  knitr::kable()
+                      "ACR"))
+  # knitr::kable()
 ```
@@
 
+```{r UDBP vs ACR time}
+
+ds %>% 
+  line_plot("")
+```
+
 ```{r correlation}
@@
   ageBase = ifelse(fVN == "Baseline", Age, NA),
-  fDM = relevel(as.factor(DM), "1")
+  fDM = relevel(as.factor(DM), "0")
   # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
@@
   dplyr::arrange(SID, VN) %>%
-  mason_gee(
-  yvars = c("ACR", "eGFR"),
-  xvars = "UDBP",
-  covars = c("VN", "ageBase", "fDM")
-  )
+
+  mason::design("gee") %>% 
+  mason::add_settings(
+    family = stats::gaussian(),
+    corstr = "ar1",
+    cluster.id = "SID"
+  ) %>% 
+  mason::add_variables("yvars", c("ACR", "eGFR")) %>% 
+  mason::add_variables("xvars", "UDBP") %>% 
+  mason::add_variables("covariates", c("VN", "ageBase", "fDM")) %>% 
+  mason::construct() %>% 
+  mason::scrub()
 ```

### 2017-03-13 doc/seminar2017.Rmd
Subject characterization table using complete data

@@
 # 
-# dsComplete <- ds %>% 
-#   dplyr::group_by(SID) %>%
-#   dplyr::filter(n() == 3) %>%
-#   dplyr::ungroup()
+dsComplete <- ds %>%
+  dplyr::group_by(SID) %>%
+  dplyr::filter(n() == 3) %>%
+  dplyr::ungroup()
 ```
@@
   strata = c("fVN"),
-  data = ds,
+  data = dsComplete,
   factorVars = c("Sex", "SmokeCigs")
@@
                       "UDBP",
+                      "UrineMicroalbumin", 
+                      "UrineCreatinine", 
                       "ACR"))

### 2017-03-20 doc/seminar2017.Rmd

@@
 
+dsTableOne <- ds %>% 
+  dplyr::filter(Sex == "Female")
+
 subCharTable <- tableone::CreateTableOne(
@@
            "Sex",
-           "Ethnicity",
-           "BMI",
-           "Waist",
            "eGFR",
            "ACR",
-           "UrineCreatinine",
-           "UrineMicroalbumin",
            "UDBP",
-           "Creatinine",
            "VitaminD",
-           "PTH",
-           "ALT",
-           "Systolic",
-           "Diastolic",
-           "MeanArtPressure",
-           "Glucose0",
-           "Glucose120",
            "dmStatus",
-           "SmokeCigs"),
+           "BirthControl",
+           "PeriodsStopped"),
   strata = c("fVN"),
-  data = dsComplete,
-  factorVars = c("Sex", "SmokeCigs")
+  data = dsTableOne,
+  factorVars = c("SmokeCigs",
+                 "BirthControl",
+                 "PeriodsStopped")
 ) %>% 
-  print(nonnormal = c("Glucose0",
-                      "Glucose120",
-                      "ALT",
-                      "UDBP",
-                      "UrineMicroalbumin", 
-                      "UrineCreatinine", 
-                      "ACR"))
-  # knitr::kable()
+  print(nonnormal = c("UDBP",
+                      "ACR")) %>% 
+  knitr::kable()
 ```

### 2017-03-21 doc/seminar2017.Rmd
Used regression for continuous variables

@@
             "Albuminuria",
-            "log uVDBP:Creatinine")
+            "log uVDBP:Creatinine (ug/mmol)")
 
@@
   scatter_plot("log(ACR)", "log(udbpCrRatio)",
-               "log Albumin:Creatinine Ratio",
-               "log UDBP:Creatinine",
+               "log Albumin:Creatinine (mg/mmol)",
+               "log uVDBP:Creatinine (ug/mmol)",
                line = TRUE)
@@
             "Estimated GFR (ml/min/1.73m^2)",
-            "log uVDBP:Creatinine")
+            "log uVDBP:Creatinine (ug/mmol)")
 
@@
 ds %>% 
-  dplyr::filter(!(eGFR_status == "Moderate")) %>% 
-  dplyr::filter(eGFR < 150) %>% 
+  # dplyr::filter(eGFR < 150) %>% 
+  dplyr::filter(fVN == "Baseline") %>% 
   dplyr::select(eGFR, udbpCrRatio) %>% 
@@
                "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
-               "log UDBP:Creatinine") +
-  ggplot2::geom_smooth(method = lm, colour = "grey50")
+               "log UDBP:Creatinine (ug/mmol)")
 
@@
 ds %>% 
-  filter(!(acrStatus == "Macroalbuminuria")) %>% 
-  cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")
+  # filter(!(acrStatus == "Macroalbuminuria")) %>% 
+  dplyr::filter(fVN == "Baseline") %>% 
+  cor.test(formula = ~ eGFR + log(udbpCrRatio), data = ., method = "spearman")
+
+```
+
+```{r glm}
+
+# Multiple Regression -------------------------------------------
+
+ds %>%
+  dplyr::filter(fVN == "Baseline") %>% 
+  # dplyr::mutate(UDBP = UDBP/1000) %>%
+  dplyr::arrange(SID, VN) %>%
+
+  mason::design("glm") %>% 
+  mason::add_settings(family = stats::gaussian()) %>% 
+  mason::add_variables("yvars", "UDBP") %>% 
+  mason::add_variables("xvars", c("ACR", "eGFR")) %>% 
+  mason::construct() %>% 
+  # mason::add_variables("covariates", c("DM", "MedsBloodPressure")) %>% 
+  # mason::construct() %>% 
+  mason::add_variables("yvars", "VitaminD") %>% 
+  mason::add_variables("xvars", "UDBP") %>% 
+  mason::construct() %>% 
+  mason::scrub()
+
 ```
@@
            facet = TRUE)
-```
-
-## Glycemic Status
-
-```{r DM boxplot}
-
-# Boxplot of uVDBP concentrations across DM categories
-
-ds %>% 
-  filter(fVN == "Baseline") %>% 
-  select(dmStatus, udbpCrRatio) %>% 
-  na.omit() %>% 
-  box_plot_slides("dmStatus", "log(udbpCrRatio)", 
-            "Glycemic Status",
-            "log uVDBP:Creatinine")
-
-# ANOVA
-
-anova <- aov(formula = log(udbpCrRatio)~eGFR_status, data = ds1)
-summary(anova)
-TukeyHSD(anova)
-rm(anova)
-```
-
-## Vitamin D
-
-```{r VitD scatterplot}
-
-ds %>% 
-  dplyr::filter(fVN == "Baseline") %>% 
-  dplyr::select(VitaminD, udbpCrRatio) %>% 
-  scatter_plot("log(udbpCrRatio)", "VitaminD",
-               "log UDBP:Creatinine",
-               "Serum 25(OH)D",
-               line = TRUE)
-
-# Spearman Correlation ------------------------------------------
-
-ds %>% 
-  dplyr::filter(fVN == "Baseline") %>% 
-  cor.test(formula = ~ udbpCrRatio + VitaminD, data = ., method = "spearman")
 ```

### 2017-03-22 R/plots.R
Improved visibility of graphs

@@ -87,8 +87,8 @@ box_plot_slides <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 12, family = "Open Sans"),
-                   axis.text.y = ggplot2::element_text(colour = "grey"),
+                   text = ggplot2::element_text(size = 14, family = "Open Sans"),
+                   axis.text.y = ggplot2::element_text(colour = "grey30"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "grey30"),
                    axis.title = ggplot2::element_text(family = "Dosis")) + #angle = 45
@@ -126,10 +126,10 @@ scatter_plot = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet =
     ggplot2::theme_minimal() +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 12, family = "Open Sans"),
-                   axis.text.y = ggplot2::element_text(colour = "grey"),
+                   text = ggplot2::element_text(size = 14, family = "Open Sans"),
+                   axis.text.y = ggplot2::element_text(colour = "grey30"), #"grey" normally
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "grey50"), #angle = 45
+                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
                    axis.title = ggplot2::element_text(family = "Dosis")) + #size = 10 normally
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)

### 2017-03-22 doc/seminar2017.Rmd
Fixed plot axis label

@@
 set_options()
+extrafont::loadfonts(device="win")
 ```
@@
                "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
-               "log UDBP:Creatinine (ug/mmol)")
+               "log uVDBP:Creatinine (ug/mmol)")
 

### 2017-03-23 doc/seminar2017.Rmd
Assessed vitamin D and inflammation (CRP)

@@
 
+## Inflammation
+
+```{r CRP}
+
+ds %>% 
+  dplyr::filter(fVN == "Baseline") %>% 
+  scatter_plot("log(CRP)", "VitaminD",
+               "CRP", "25(OH)D")
+
+ds %>%
+  dplyr::filter(fVN == "Baseline") %>% 
+  # dplyr::mutate(UDBP = UDBP/1000) %>%
+  dplyr::arrange(SID, VN) %>%
+
+  mason::design("glm") %>% 
+  mason::add_settings(family = stats::gaussian()) %>% 
+  mason::add_variables("yvars", "VitaminD") %>% 
+  mason::add_variables("xvars", "CRP") %>% 
+  mason::construct() %>% 
+  mason::scrub()
+```
+
+## Hyperfiltration
+
 ```{r hyperfiltration}

### 2017-03-23 R/fetch_data.R
Added CRP as variable

@@ -109,6 +109,7 @@ fetch_data <- function() {
       UrinaryCalcium,
       fMedsBP,
       SmokeCigs,
+      CRP,
       Canoe,
       BirthControl,
       PeriodsStopped,

### 2017-04-26 R/plots.R
line graph shows all UDBP measures

@@ -205,13 +205,11 @@ histo_plot = function(data, variable, bin, xlab='', facet = FALSE) {
 #'   dplyr::filter(SID < 1050 ) %>%
 #'   line_plot("fVN", "UDBP", "SID", "Visit Number", "UDBP")
 line_plot = function(data, xvar, yvar, byvar, xlab='', ylab='') {
-  ggplot2::ggplot(data, ggplot2::aes_string(x = xvar, y = yvar,
-                                            group = byvar, colour = byvar)) +
+  ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar,
+                                            group=byvar, colour=byvar)) +
     ggplot2::geom_line() +
     ggplot2::geom_point() +
     ggplot2::theme_minimal() +
-    ggplot2::scale_fill_brewer(palette = "Dark2") +
-    ggplot2::scale_color_brewer(palette = "Dark2") +
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),

### 2017-04-27 R/plots.R
Changed alpha for line plot to make it more aesthetic

@@ -207,8 +207,8 @@ histo_plot = function(data, variable, bin, xlab='', facet = FALSE) {
 line_plot = function(data, xvar, yvar, byvar, xlab='', ylab='') {
   ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar,
                                             group=byvar, colour=byvar)) +
-    ggplot2::geom_line() +
-    ggplot2::geom_point() +
+    ggplot2::geom_line(alpha = 0.5) +
+    # ggplot2::geom_point(alpha = 0.5) +
     ggplot2::theme_minimal() +
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),

### 2017-04-28 R/plots.R
new colour theme for ada slides (boxplot)

@@ -104,6 +104,53 @@ box_plot_slides <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
 }
 
 
+#' Boxplots where the colour theme matches ada poster
+#'
+#' @param data
+#' @param xvar
+#' @param yvar
+#' @param xlab
+#' @param ylab
+#' @param facet
+#'
+#' @return Plot
+#'
+#' @examples
+box_plot_poster <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
+  myboxplot <- ggplot2::ggplot(data, ggplot2::aes_string(x = xvar, y = yvar)) +
+    ggplot2::geom_boxplot(ggplot2::aes_string(colour = xvar, fill = xvar),
+                          outlier.shape = NA) +
+    ggplot2::stat_summary(geom = "crossbar", width = 0.65, fatten = 0, color = "white",
+                          fun.data = function(x){
+                            return(c(y = median(x), ymin = median(x), ymax = median(x)))
+                          }) +
+    ggplot2::geom_jitter(ggplot2::aes_string(),
+                         position = ggplot2::position_jitter(width = 0.15, height = 0),
+                         alpha = 0.3,
+                         size = 1,
+                         colour = "grey30") +
+    ggplot2::theme_minimal() +
+    ggplot2::scale_color_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
+    ggplot2::scale_fill_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
+    ggplot2::theme(legend.position = "none",
+                   panel.grid.major.x = ggplot2::element_blank(),
+                   axis.line.y = ggplot2::element_blank(),
+                   text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
+                   axis.text.y = ggplot2::element_text(colour = "grey30"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
+                   axis.text.x = ggplot2::element_text(colour = "grey30"),
+                   axis.title = ggplot2::element_text(family = "Montserrat Medium")) + #angle = 45
+    ggplot2::xlab(xlab) +
+    ggplot2::ylab(ylab)
+
+  if(facet == TRUE) {
+    myboxplot +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else {
+    myboxplot
+  }
+}
+
 # Scatterplot -------------------------------------------------------------
 
 

### 2017-05-23 R/plots.R
scatter plot function with ada poster theme

@@ -195,7 +195,46 @@ scatter_plot = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet =
   }
 }
 
-
+#' Scatterplot where the colour theme matches ada poster
+#'
+#' @param data Clearned data
+#' @param xvar Variable on x-axis
+#' @param yvar Variable on y-axis
+#' @param xlab x-axis label
+#' @param ylab y-axis label
+#' @param line Line of best fit
+#' @param facet Split the plot by visit number
+#'
+#' @return Plot
+#'
+#' @examples
+scatter_plot_poster = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
+  myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
+    ggplot2::geom_point(colour = "#adc9b7", size = 1) + #mapping=aes(color=mcr_status)
+    ggplot2::theme_minimal() +
+    ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
+                   axis.line.y = ggplot2::element_blank(),
+                   text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
+                   axis.text.y = ggplot2::element_text(colour = "grey30"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
+                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
+                   axis.title = ggplot2::element_text(family = "Montserrat Medium")) +
+    ggplot2::xlab(xlab) +
+    ggplot2::ylab(ylab)
+  if(line == TRUE & facet == FALSE) {
+    myplot +
+      ggplot2::geom_smooth(method = loess, colour = "grey50")
+  } else if(line == TRUE & facet == TRUE) {
+    myplot +
+      ggplot2::geom_smooth(method = loess, colour = "grey50") +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else if(line == FALSE & facet == TRUE) {
+    myplot +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else {
+    myplot
+  }
+}
 
 # Histogram ---------------------------------------------------------------
 

### 2017-05-23 doc/adaAbstract.Rmd
plots for ada poster

@@
   na.omit() %>%
-  box_plot("acrStatus", "log(udbpCrRatio)", 
+  box_plot_poster("acrStatus", "log(udbpCrRatio)", 
             "Albuminuria",
@@
   filter(fVN == "Baseline") %>% 
-  select(ACR, udbpCrRatio) %>% 
-  na.omit() %>% 
-  scatter_plot("log(ACR)", "log(udbpCrRatio)",
+  scatter_plot_poster("log(ACR)", "log(udbpCrRatio)",
                "log Albumin:Creatinine Ratio",
-               "log UDBP:Creatinine",
-               line = TRUE)
+               "log UDBP:Creatinine")
 
@@
   filter(fVN == "Baseline") %>% 
-  select(eGFR_status, udbpCrRatio) %>% 
+  select(eGFRStatus, udbpCrRatio) %>% 
   na.omit() %>% 
-  box_plot2("eGFR_status", "log(udbpCrRatio)", 
+  box_plot_poster("eGFRStatus", "log(udbpCrRatio)", 
             "Estimated GFR (ml/min/1.73m^2)",
@@
 ds %>% 
-  dplyr::filter(!(eGFR_status == "Moderate")) %>% 
-  dplyr::filter(eGFR < 150) %>% 
-  dplyr::select(eGFR, udbpCrRatio) %>% 
-  na.omit() %>% 
-  scatter_plot("eGFR", "log(udbpCrRatio)",
+  dplyr::filter(fVN == "Baseline") %>% 
+  scatter_plot_poster("eGFR", "log(udbpCrRatio)",
                "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
-               "log UDBP:Creatinine") +
-  ggplot2::geom_smooth(method = lm, colour = "grey50")
+               "log UDBP:Creatinine")
 

### 2017-05-23 doc/adaAbstract.Rmd
gee model for ada abstract with new covariates

@@
 
-```{r mason}
+```{r gee}
 
@@
   ageBase = ifelse(fVN == "Baseline", Age, NA),
-  fDM = relevel(as.factor(DM), "1")
-  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
+  DM = ifelse(DM == 1, "diabetes", "non_dia"),
+  fDM = relevel(as.factor(DM), "non_dia"),
+  Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
+  Ethnicity = relevel(as.factor(Ethnicity), "Other")
   ) %>%
@@
   dplyr::arrange(SID, VN) %>%
-  mason_gee(
-  yvars = c("ACR", "eGFR"),
-  xvars = "UDBP",
-  covars = c("VN", "ageBase", "fDM")
-  )
+  mason::design("gee") %>% 
+  mason::add_settings(
+      family = stats::gaussian(),
+      corstr = 'ar1',
+      cluster.id = 'SID'
+    ) %>%
+  mason::add_variables("yvars", c("ACR", "eGFR")) %>% 
+  mason::add_variables("xvars", "UDBP") %>% 
+  mason::add_variables("covariates", 
+                       c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM")) %>% 
+  mason::construct() %>%
+  mason::scrub()
 ```

### 2017-05-24 R/plots.R
plot GEE results using seer package (lwjohnst86)

@@ -306,3 +306,20 @@ line_plot = function(data, xvar, yvar, byvar, xlab='', ylab='') {
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 }
+
+
+# GEE Plot ----------------------------------------------------------------
+
+#' GEE Plot
+#'
+#' @return
+#' @export
+#'
+#' @examples
+gee_plot = function(gee_results, xlab = "") {
+  seer::view_main_effect(gee_results, "dot.size", group.label.switch = 'y') +
+    ggplot2::theme(legend.position = 'right') +
+    ggplot2::theme_bw() +
+    ggplot2::ylab('') +
+    ggplot2::xlab(xlab)
+}

### 2017-05-24 doc/adaAbstract.Rmd
gee plot

@@
 
-```{r}
-gee3 <- readRDS(file = "../data/ds.Rds") %>%
-  dplyr::mutate(
-  udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
-  ageBase = ifelse(fVN == "Baseline", Age, NA),
-  VitaminD3 = ifelse(fVN == "6Year", NA, VitaminD)
-  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
-  ) %>%
-  dplyr::arrange(SID, fVN) %>%
-  dplyr::group_by(SID) %>%
-  tidyr::fill(udbpBase, ageBase) %>%
-  dplyr::ungroup() %>%
-  # dplyr::group_by(VN) %>%
-  # dplyr::mutate(udbpBase = as.numeric(scale(udbpBase))) %>%
-  dplyr::mutate(UDBP = UDBP/1000) %>%
-  # dplyr::ungroup() %>% 
-  dplyr::arrange(SID, VN) %>%
-  mason_gee(
-  yvars = c("VitaminD3"),
-  xvars = "UDBP",
-  covars = c("VN", "ageBase", "DM")
-  )
+```{r gee plot}
+
+gee %>% 
+  dplyr::mutate(Xterms = term) %>% 
+  dplyr::filter(!term == "(Intercept)") %>% 
+  dplyr::mutate(Yterms = factor(Yterms, 
+                                levels = c("ACR", "eGFR"),
+                                ordered = TRUE),
+                Xterms = factor(Xterms,
+                                levels = rev(c("<-Xterm",
+                                           "VN",
+                                           "ageBase",
+                                           "SexMale",
+                                           "EthnicityEuropean",
+                                           "BMI",
+                                           "fDMdiabetes")),
+                                labels = rev(c("uVDBP",
+                                           "Follow-up Duration",
+                                           "Baseline Age",
+                                           "Sex (male)",
+                                           "Ethnicity (European)",
+                                           "BMI",
+                                           "Diabetes")),
+                                ordered = TRUE)) %>% 
+  arrange(Xterms) %>% 
+  gee_plot(xlab = "Unit change for every unit increase in uVDBP")
 ```

### 2017-05-29 R/plots.R
added grey borders to poster themes

@@ -129,17 +129,18 @@ box_plot_poster <- function(data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
                          alpha = 0.3,
                          size = 1,
                          colour = "grey30") +
-    ggplot2::theme_minimal() +
     ggplot2::scale_color_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
     ggplot2::scale_fill_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
+    ggplot2::theme_grey() +
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
+                   plot.background = ggplot2::element_rect(colour = "grey60"),
                    text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
                    axis.text.y = ggplot2::element_text(colour = "grey30"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "grey30"),
-                   axis.title = ggplot2::element_text(family = "Montserrat Medium")) + #angle = 45
+                   axis.title = ggplot2::element_text(family = "Montserrat Medium")) +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
@@ -210,9 +211,10 @@ scatter_plot = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet =
 #' @examples
 scatter_plot_poster = function(data, xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
   myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
-    ggplot2::geom_point(colour = "#adc9b7", size = 1) + #mapping=aes(color=mcr_status)
-    ggplot2::theme_minimal() +
+    ggplot2::geom_point(colour = "#adc9b7", size = 1) +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
+                   panel.grid.minor.x = ggplot2::element_blank(),
+                   plot.background = ggplot2::element_rect(colour = "grey60"),
                    axis.line.y = ggplot2::element_blank(),
                    text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
                    axis.text.y = ggplot2::element_text(colour = "grey30"),
[…]

### 2017-05-29 doc/adaAbstract.Rmd
removed gee graph that is non-functional

@@
 
-# Boxplot of ACR in uVDBP categories
-dsBase %>% 
-  select(UDBPStatus, ACR) %>% 
-  na.omit() %>% 
-  box_plot("UDBPStatus", "log(ACR)",
-           "Urinary VDBP status",
-           "log Urine ACR (mg/mmol)")
-
 # ANOVA
@@
 ```{r hyperfiltration}
+
 ds %>% 
@@
   arrange(Xterms) %>% 
-  gee_plot(xlab = "Unit change for every unit increase in uVDBP")
+  gee_plot(xlab = "Percent difference in outcome for every unit increase in uVDBP")
 ```
-
-Graph
-```{r gee plot}
-
-gee %>%
-  dplyr::mutate(p.value = ifelse(p.value > 0.05, 1, 0.04)) %>%
-  seer::view_main_effect(
-  graph.options = 'dot.size',
-  groups = '~Yterms',
-  legend.title = 'p-values',
-  xlab = 'Percent difference with 95% CI in the outcomes\nfor each unit increase in baseline uVDBP',
-  ylab = 'Kidney Measures'
-  ) +
-  graph_theme(ticks = FALSE, legend.pos = 'right') +
-  ggplot2::theme(legend.margin = grid::unit(0, 'cm'))
-
-```

### 2017-06-01 doc/adaAbstract.Rmd
standard deviation output for gee instead of unit change

@@
 
-```{r correlation}
-dsGEE <- ds %>% 
-  dplyr::mutate(udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
-                ageBase = ifelse(fVN == "Baseline", Age, NA),
-                bmiBase = ifelse(fVN == "Baseline", BMI, NA),
-                waistBase = ifelse(fVN == "Baseline", Waist, NA), 
-                Sex = ifelse(Sex == "Male", 1, 0),
-                Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
-                Ethnicity = ifelse(Ethnicity == "European", 0, 1),
-                fMedsBP = ifelse(fMedsBP == "Yes", 1, 0)
-                # vitdBase = ifelse(fVN == "6Year", NA, VitaminD)
-                ) %>% 
-  dplyr::arrange(SID, fVN) %>% 
-  dplyr::group_by(SID) %>% 
-  tidyr::fill(udbpBase, ageBase, bmiBase, waistBase) %>% 
-  dplyr::ungroup() %>%
-  dplyr::arrange(SID, VN) %>% 
-  dplyr::select(udbpBase, ACR, eGFR, VN, ageBase, Sex, Ethnicity, BMI, 
-                Waist, DM, fMedsBP)
-
-round(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"), 2)
-
-# Heatmap
-
-corrplot::corrplot(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"),
-                   method = "color")
-
-```
-
 ```{r gee}
@@
   dplyr::ungroup() %>%
-  dplyr::mutate(UDBP = UDBP/1000) %>%
+  # dplyr::mutate(UDBP = UDBP/1000) %>%
   dplyr::arrange(SID, VN) %>%
@@
   mason::add_variables("xvars", "UDBP") %>% 
-  mason::add_variables("covariates", 
-                       c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM")) %>% 
+  mason::add_variables("covariates", c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "fDM")) %>% 
   mason::construct() %>%
-  mason::scrub()
+  mason::scrub() %>% 
+  mason::polish_transform_estimates(function(x) (exp(x) - 1) * 100) %>%
+  dplyr::select(Yterms, Xterms, term, estimate, conf.low, conf.high, p.value)
 ```
@@
   arrange(Xterms) %>% 
-  gee_plot(xlab = "Percent difference in outcome for every unit increase in uVDBP")
+  gee_plot(xlab = "Percent difference with 95% CI in outcome for every SD increase in uVDBP and covariates")
 ```

### 2017-06-01 R/plots.R

@@ -336,4 +336,3 @@ gee_plot = function(gee_results, xlab = "") {
     ggplot2::ylab('') +
     ggplot2::xlab(xlab)
 }
-# "#2b3443", "#adc9b7", "#ffd91e"

### 2017-06-02 doc/adaAbstract.Rmd
fixed scaling problems with gee output

@@
 ```{r setup, include=FALSE}
+
+# poster size: print at 63" x 31.5"
+
 knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
@@
   dplyr::ungroup() %>%
-  # dplyr::mutate(UDBP = UDBP/1000) %>%
+  dplyr::mutate(UDBP = UDBP/1000) %>%
   dplyr::arrange(SID, VN) %>%
@@
   mason::scrub() %>% 
-  mason::polish_transform_estimates(function(x) (exp(x) - 1) * 100) %>%
   dplyr::select(Yterms, Xterms, term, estimate, conf.low, conf.high, p.value)
@@
                                            "fDMdiabetes")),
-                                labels = rev(c("uVDBP",
-                                           "Follow-up Duration",
-                                           "Baseline Age",
+                                labels = rev(c("uVDBP (ug/mL)",
+                                           "Follow-up Duration (Years)",
+                                           "Baseline Age (Years)",
                                            "Sex (male)",
                                            "Ethnicity (European)",
-                                           "BMI",
+                                           "BMI (kg/m^2)",
                                            "Diabetes")),
@@
   arrange(Xterms) %>% 
-  gee_plot(xlab = "Percent difference with 95% CI in outcome for every SD increase in uVDBP and covariates")
+  gee_plot(xlab = "Unit difference with 95% CI in outcome for every unit increase in uVDBP and covariates")
 ```

### 2017-07-27 R/fetch_data.R
table 1 (subject characteristics) completed

@@ -69,10 +69,11 @@ fetch_data <- function() {
                        breaks = c(0, 1.23, 60, Inf),
                        labels = c("Trace", "Normal", "High"),
                        ordered_result = TRUE),
-      UDBPquartile = cut(UDBP,
-                         breaks = c(0, 12.015, 45.25, 90.975, 2123),
-                         labels = c("1st", "2nd", "3rd", "4th"),
-                         ordered_result = TRUE),
+      UDBPQuartile = ntile(UDBP, 4),
+      # UDBPQuartile = cut(UDBP,
+      #                    breaks = c(0, 12.015, 45.25, 90.975, 2123),
+      #                    labels = c("1st", "2nd", "3rd", "4th"),
+      #                    ordered_result = TRUE),
       udbpCrRatio = UDBP / UrineCreatinine,
       logUDBPRatio = log(udbpCrRatio)
     ) %>%
@@ -97,6 +98,7 @@ fetch_data <- function() {
       acrStatus,
       eGFRStatus,
       UDBPStatus,
+      UDBPQuartile,
       eGFR,
       ACR,
       UrineMicroalbumin,
@@ -119,8 +121,10 @@ fetch_data <- function() {
       BirthControl,
       PeriodsStopped,
       PeriodsStoppedAge,
+      MedsBloodPressure,
+      MedsDxDM
       # MedName,
-      dplyr::matches("meds")
+      # dplyr::matches("meds")
     )
 
   # Final dataset object

### 2017-08-08 index.Rmd
initial commit

---
title: "Urinary Vitamin D Binding Protein Loss as a Potential Biomarker for Nephropathy in Patients at Risk for Type 2 Diabetes: The Prospective Metabolism and Islet Cell Evaluation (PROMISE) Cohort Study"
author: "Windy Wang"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib] #packages.bib
biblio-style: apalike
link-citations: yes
github-repo: windyzn/master-thesis
description: "An exploration of the relationships between urinary vitamin D binding protein and kidney dysfunction, glycemic status, and 25(OH)D levels in subjects at risk for type 2 diabetes."
---

# Preface {-}

This is my master's thesis written using bookdown.

# Acknowledgements {-}

Thank you for your help.

# Abstract {-}

**Background**: Recent studies have reported elevated urinary vitamin D binding protein (uVDBP) in patients with type 1 diabetes, although the utility of uVDBP to predict declines in kidney function over time has not been examined.  Our objective was to assess the association of uVDBP with longitudinal declines in kidney function.

**Methods**: Data were derived from PROMISE, a cohort of adults at risk for type 2 diabetes. We used data from 3 clinical visits over 6 years (n = 730). Urinary albumin-to-creatinine ratio (ACR) and eGFR were used as markers of kidney function. Measurements of uVDBP were performed with ELISA and normalized to urine creatinine. Generalized estimating equations (GEE) evaluated longitudinal associations of uVDBP with kidney measures. Covariates included visit number, baseline age, and glycemic status.

**Results**: Renal VDBP loss increased with ACR severity at baseline. Subjects with normoalbuminuria, microalbuminuria, and macroalbuminuria had median log uVDBP concentrations of 1.62 (IQR 0.95 – 1.98), 2.63 (IQR 2.25 – 3.37), and 2.48 (IQR 2.19 – 2.77) respectively (all p < 0.001), and ACR positively correlated with uVDBP concentrations (r = 0.38, p < 0.001). There was no significant association between uVDBP and eGFR at baseline (p = 0.20).  Adjusted GEE models indicated that each unit difference in uVDBP (μg/mL) predicted higher ACR (β = 0.55, p < 0.001) and lower eGFR (β = -0.50, p = 0.013) over 6 years.

**Significance**: These results suggest that loss of uVDBP over time may be a useful marker for predicting renal dysfunction in subjects at risk for diabetes.

The code can be found at my [GitHub repository](https://github.com/windyzn/master-thesis)

### 2017-08-08 R/gee.R
set default parameters for data variable

@@ -40,7 +40,7 @@ prep_gee_data <- function(data) {
 #' @export
 #'
 #' @examples
-analyze_gee <- function(data,
+analyze_gee <- function(data = project_data,
                         y = outcomes,
                         x = xvar,
                         covars = covariates,

### 2017-08-08 R/interaction.R
plot for interaction

@@ -5,6 +5,80 @@
 #'
 #' @examples
 #'
-interaction <- function() {
+plot_interaction <- function(data, y, x,
+                             fatty.acid.levels = c(2, 3, 4),
+                             covars = covariates,
+                             title = 'A',
+                             interaction.term = 'VN',
+                             is.fatty.acid = TRUE,
+                             y.axis.label = NULL) {
+  fa.lvl <- fatty.acid.levels
+  if (!fa.lvl %in% c(2, 3, 4)) {
+    print('Please select either 2, 3, or 4 for the fatty.acid.levels')
+  } else if (fa.lvl == 2) {
+    lvls <- 0:2/2
+    lvl.labs <- c('<50%', '>50%')
+  } else if (fa.lvl == 3) {
+    lvls <- 0:3/3
+    lvl.labs <- c('<33%', '33-66%', '>66%')
+  } else if (fa.lvl == 4) {
+    lvls <- 0:4/4
+    lvl.labs <- c('Lowest', 'Low', 'High', 'Highest')
+  }
 
+  if (is.null(y.axis.label)) {
+    y.axis.label <- paste0('log(', substring(y, 2), ')')
+  }
+
+  fatty.acid.name <- renaming_fats(x) %>%
+    gsub('(\\D\\D)\\.', '\\U\\1 ', ., perl = TRUE)
+
+  if (is.fatty.acid) {
+    fatty.acid.name <- paste(fatty.acid.name,
+                             ifelse(grepl('pct_', x), '(mol%)', '\n(nmol/mL)'))
+  }
+
+  geeFormula <- paste(y, 'f.fat', sep = ' ~ ') %>%
+    paste(., paste(covars, collapse = ' + '),
+          paste('f.fat', interaction.term, sep = ':'), sep = ' + ') %>%
+    as.formula()
+
+  prep.fit <- data %>%
+    dplyr::select_(.dots = c(y, x, covars, interaction.term, 'SID')) %>%
+    dplyr::filter(complete.cases(.)) %>%
+    dplyr::rename_('fa' = x, 'VN' = interaction.term) %>%
+    dplyr::ungroup() %>%
+    dplyr::mutate(f.fat = cut(fa, stats::quantile(fa, probs = lvls),
+                              include.lowest=TRUE, labels = lvl.labs),
+                  VN = factor(VN, ordered = TRUE))
+
+  fit <-
+    geepack::geeglm(geeFormula, id = SID, family = stats::gaussian,
[…]

### 2017-08-08 R/interaction.R
removed unnecessary luke code

@@ -6,7 +6,6 @@
 #' @examples
 #'
 plot_interaction <- function(data, y, x,
-                             fatty.acid.levels = c(2, 3, 4),
                              covars = covariates,
                              title = 'A',
                              interaction.term = 'VN',

### 2017-08-08 R/plots.R
refined line plot

@@ -297,11 +297,10 @@ histo_plot = function(data=project_data, variable, bin, xlab='', facet = FALSE) 
 line_plot = function(data=project_data, xvar, yvar, byvar, xlab='', ylab='') {
   ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar,
                                             group=byvar, colour=byvar)) +
-    ggplot2::geom_line(alpha = 0.5) +
-    # ggplot2::geom_point(alpha = 0.5) +
+    ggplot2::geom_line(alpha = 0.7) +
+    ggplot2::geom_point(alpha = 0.5) +
     ggplot2::theme_minimal() +
-    ggplot2::theme(legend.position = "none",
-                   panel.grid.major.x = ggplot2::element_blank(),
+    ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
                    axis.text.y = ggplot2::element_text(colour = "grey"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
@@ -311,6 +310,7 @@ line_plot = function(data=project_data, xvar, yvar, byvar, xlab='', ylab='') {
     ggplot2::ylab(ylab)
 }
 
+#legend.position = "none" was the first one under theme
 
 # GEE Plot ----------------------------------------------------------------
 

### 2017-08-09 R/progression.R
initial commit

@@ -1,0 +1,32 @@
+
+summarise_progress_data <- function(data, by) {
+  data %>%
+    dplyr::group_by(by, fVN) %>%
+    dplyr::summarise(UDBP=mean(log(UDBP)))
+}
+
+#' Plot progression of UDBP over time (visit number) grouped by variables
+#'
+#' @param data
+#' @param summary_data
+#' @param xvar
+#' @param yvar
+#' @param byvar
+#' @param groupby
+#'
+#' @return
+#' @export
+#'
+#' @examples
+plot_progress_data <- function(data, summary_data,
+                               xvar = "fVN", yvar = "log(UDBP)",
+                               byvar, groupby = "SID") {
+  data %>%
+    ggplot2::ggplot(ggplot2::aes_string(x = xvar,
+                               y = yvar,
+                               colour = byvar)) +
+    ggplot2::geom_line(ggplot2::aes_string(group = groupby), alpha = 0.1) +
+    ggplot2::geom_line(data = summary_data, ggplot2::aes_string(group = byvar),
+                       alpha = 0.7, size = 2) +
+    ggplot2::ylab("log(UDBP)")
+}

### 2017-08-09 R/plots.R
removed font specifications for basic scatterplot function

@@ -175,19 +175,19 @@ scatter_plot = function(data=project_data,
     ggplot2::theme_minimal() +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 14, family = "Open Sans"),
+                   text = ggplot2::element_text(size = 14), #family = "Open Sans"
                    axis.text.y = ggplot2::element_text(colour = "grey30"), #"grey" normally
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
-                   axis.title = ggplot2::element_text(family = "Dosis")) + #size = 10 normally
+                   axis.text.x = ggplot2::element_text(colour = "grey30")) + #angle = 45
+                   # axis.title = ggplot2::element_text(family = "Dosis")) + #size = 10 normally
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
   if(line == TRUE & facet == FALSE) {
     myplot +
-      ggplot2::geom_smooth(method = lm, colour = "grey50")
+      ggplot2::geom_smooth(method = loess, colour = "grey50")
   } else if(line == TRUE & facet == TRUE) {
     myplot +
-      ggplot2::geom_smooth(method = lm, colour = "grey50") +
+      ggplot2::geom_smooth(method = loess, colour = "grey50") +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else if(line == FALSE & facet == TRUE) {
     myplot +

### 2017-08-09 R/summary.R
removed summary function as it was not necessary with new tidyverse updates

@@ -1,42 +1,0 @@
-## Summarizes data.
-## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   groupvars: a vector containing names of columns that contain grouping variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
-                      conf.interval=.95, .drop=TRUE) {
-  library(plyr)
-
-  # New version of length which can handle NA's: if na.rm==T, don't count them
-  length2 <- function (x, na.rm=FALSE) {
-    if (na.rm) sum(!is.na(x))
-    else       length(x)
-  }
-
-  # This does the summary. For each group's data frame, return a vector with
-  # N, mean, and sd
-  datac <- ddply(data, groupvars, .drop=.drop,
-                 .fun = function(xx, col) {
-                   c(N    = length2(xx[[col]], na.rm=na.rm),
-                     mean = mean   (xx[[col]], na.rm=na.rm),
-                     sd   = sd     (xx[[col]], na.rm=na.rm)
-                   )
-                 },
-                 measurevar
-  )
-
-  # Rename the "mean" column
-  datac <- rename(datac, c("mean" = measurevar))
-
-  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
-
-  # Confidence interval multiplier for standard error
-  # Calculate t-statistic for confidence interval:
-  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
-  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
-  datac$ci <- datac$se * ciMult
-
-  return(datac)
-}

### 2017-08-14 R/interaction.R
removed export

@@ -1,7 +1,6 @@
 #' Interaction with time
 #'
-#' @return
-#' @export
+#' @return Plot
 #'
 #' @examples
 #'

### 2017-08-14 R/plots.R
improved aesthetics of scatter_plot function to be more professional

@@ -170,18 +170,23 @@ box_plot_poster <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
 #' @examples
 scatter_plot = function(data=project_data,
                         xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
+
   myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
-    ggplot2::geom_point(colour = "#0db7c4", size = 1) + #mapping=aes(color=mcr_status)
+    ggplot2::geom_point(size = 1.5, alpha = 0.8) + #mapping=aes(color=mcr_status), #0db7c4
+    ggplot2::scale_fill_brewer(palette = "Dark2") +
+    ggplot2::scale_color_brewer(palette = "Dark2") +
     ggplot2::theme_minimal() +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
+                   panel.grid.minor.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 14), #family = "Open Sans"
+                   text = ggplot2::element_text(size = 11, family = "Open Sans"), #family = "Open Sans"
                    axis.text.y = ggplot2::element_text(colour = "grey30"), #"grey" normally
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "grey30")) + #angle = 45
                    # axis.title = ggplot2::element_text(family = "Dosis")) + #size = 10 normally
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
+
   if(line == TRUE & facet == FALSE) {
     myplot +
       ggplot2::geom_smooth(method = loess, colour = "grey50")
@@ -216,7 +221,7 @@ scatter_plot_poster = function(data=project_data,
     ggplot2::geom_point(colour = "#adc9b7", size = 1) +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    panel.grid.minor.x = ggplot2::element_blank(),
-                   plot.background = ggplot2::element_rect(colour = "grey60"),
+                   panel.background = ggplot2::element_rect(colour = "grey60"),
                    axis.line.y = ggplot2::element_blank(),
                    text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
                    axis.text.y = ggplot2::element_text(colour = "grey30"),

### 2017-08-14 R/gee.R
renamed variables

@@ -10,7 +10,7 @@
 #' @export
 
 
-prep_gee_data <- function(data) {
+prep_mason_data <- function(data) {
   data %>%
     dplyr::mutate(
       udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
@@ -40,9 +40,9 @@ prep_gee_data <- function(data) {
 #' @export
 #'
 #' @examples
-analyze_gee <- function(data = project_data,
+mason_gee <- function(data = project_data,
                         y = outcomes,
-                        x = xvar,
+                        x = predictors,
                         covars = covariates,
                         intvar = NULL) {
 

### 2017-08-14 R/regression.R
initial commit

@@ -1,0 +1,28 @@
+#' Multiple linear regression
+#'
+#' @param data
+#' @param y
+#' @param x
+#' @param covars
+#' @param intvar
+#'
+#' @return
+#'
+#' @examples
+mason_glm <- function(data = project_data,
+                        y = outcomes,
+                        x = predictors,
+                        covars = NULL) {
+
+  co <- !is.null(covars)
+
+  data %>%
+    mason::design("glm") %>%
+    mason:::add_settings(family = stats::gaussian()) %>%
+    mason::add_variables("yvars", y) %>%
+    mason::add_variables("xvars", x) %>%
+    mason::construct() %>%
+
+    mason::construct() %>%
+    mason::scrub()
+}

### 2017-08-17 R/fetch_data.R
fixed dmStatus missingness

@@ -42,6 +42,7 @@ fetch_data <- function() {
                          ifelse(IFG == 1 |
                                   IGT == 1, "PreDM",
                                 "NGT")),
+      dmStatus = ifelse(is.na(dmStatus), "NGT", dmStatus),
       dmStatus = factor(dmStatus,
                         levels = c("NGT", "PreDM", "DM"),
                         ordered = TRUE),

### 2017-08-23 R/fetch_data.R
new hypertension variable

@@ -35,9 +35,14 @@ fetch_data <- function() {
       fMedsBP = factor(
         MedsBloodPressure,
         levels = c(0, 1),
-        labels = c("No", "Yes"),
-        ordered = TRUE
+        labels = c("No", "Yes")
       ),
+      Hypertension = ifelse(!(is.na(MedHistHypertenAge)), 1,
+                            ifelse(Systolic > 140 | Diastolic > 90, 1,
+                                   0)),
+      Hypertension = factor(Hypertension,
+                            levels = c(0, 1),
+                            labels = c("No", "Yes")),
       dmStatus = ifelse(DM == 1, "DM",
                          ifelse(IFG == 1 |
                                   IGT == 1, "PreDM",
@@ -111,6 +116,7 @@ fetch_data <- function() {
       MeanArtPressure,
       Systolic,
       Diastolic,
+      Hypertension,
       PTH,
       ALT,
       UrinaryCalcium,
@@ -122,6 +128,7 @@ fetch_data <- function() {
       BirthControl,
       PeriodsStopped,
       PeriodsStoppedAge,
+      MedHistHypertenAge,
       MedsBloodPressure,
       MedsDxDM,
       MonthsFromBaseline

### 2017-08-24 R/fetch_data.R
replaced UDBP quantiles with tertiles

@@ -75,7 +75,7 @@ fetch_data <- function() {
                        breaks = c(0, 1.23, 60, Inf),
                        labels = c("Trace", "Normal", "High"),
                        ordered_result = TRUE),
-      UDBPQuartile = ntile(UDBP, 4),
+      UDBPTertile = ntile(UDBP, 3),
       # UDBPQuartile = cut(UDBP,
       #                    breaks = c(0, 12.015, 45.25, 90.975, 2123),
       #                    labels = c("1st", "2nd", "3rd", "4th"),
@@ -104,7 +104,7 @@ fetch_data <- function() {
       acrStatus,
       eGFRStatus,
       UDBPStatus,
-      UDBPQuartile,
+      UDBPTertile,
       eGFR,
       ACR,
       UrineMicroalbumin,

### 2017-08-24 R/regression.R
added covariate options

@@ -21,8 +21,13 @@ mason_glm <- function(data = project_data,
     mason:::add_settings(family = stats::gaussian()) %>%
     mason::add_variables("yvars", y) %>%
     mason::add_variables("xvars", x) %>%
-    mason::construct() %>%
-
-    mason::construct() %>%
+    mason::construct() %>% {
+      if (co) {
+        mason::add_variables("covariates", covars) %>%
+        mason::construct()
+      } else {
+        .
+      }
+    } %>%
     mason::scrub()
 }

### 2017-08-24 R/gee.R
improved documentation on prep_mason_data()

@@ -4,7 +4,9 @@
 # Grab or combine data ----------------------------------------------------
 
 
-#' Prepare the project data for analysis through GEE
+#' Prepare the project data for analysis through GEE. The output is a dataframe
+#' with baseline values. Subjects with macroalbuminuria and moderate kidney
+#' disease at baseline is omitted.
 #'
 #' @param data project data
 #' @export

### 2017-08-24 R/progression.R
progression plot that does not separate by variable

@@ -10,12 +10,13 @@
 #' @export
 #'
 #' @examples
-plot_progress_data <- function(data,
+plot_progress_by <- function(data,
                                xvar = "fVN", yvar = "log(UDBP)",
+                               ylab = "log(UDBP)",
                                byvar, groupby = "SID") {
   m <- data %>%
     dplyr::group_by_(byvar, xvar) %>%
-    dplyr::summarise(UDBP=mean(log(UDBP)))
+    dplyr::summarise(UDBP=mean(UDBP))
 
   data %>%
     ggplot2::ggplot(ggplot2::aes_string(x = xvar,
@@ -24,5 +25,24 @@ plot_progress_data <- function(data,
     ggplot2::geom_line(ggplot2::aes_string(group = groupby), alpha = 0.1) +
     ggplot2::geom_line(data = m, ggplot2::aes_string(group = byvar),
                        alpha = 0.7, size = 2) +
-    ggplot2::ylab("log(UDBP)")
+    ggplot2::ylab(ylab)
+}
+
+
+
+
+plot_progress <- function(data = project_data,
+                               xvar = "fVN", yvar,
+                               ylab,
+                               groupby = "SID") {
+  m <- data %>%
+    dplyr::group_by_(yvar, xvar) %>%
+    dplyr::summarise(yvar=mean(yvar))
+
+  data %>%
+    ggplot2::ggplot(ggplot2::aes_string(x = xvar,
+                                        y = yvar)) +
+    ggplot2::geom_line(ggplot2::aes_string(group = groupby), alpha = 0.1) +
+    ggplot2::geom_line(data = m, alpha = 0.7, size = 2) +
+    ggplot2::ylab(ylab)
 }

### 2017-08-25 R/gee.R
modified function to show values for covariates in gee

@@ -69,8 +69,9 @@ mason_gee <- function(data = project_data,
       }
     } %>%
     mason::construct() %>%
-    mason::scrub() %>%
-    mason::polish_filter(extract_term, "term")
+    mason::scrub()
+  # %>%
+  #   mason::polish_filter(extract_term, "term")
 }
 
 

### 2017-08-28 R/gee.R
function for plotting gee results

@@ -98,40 +98,30 @@ explore_gee <- function(data = project_data,
 #' @export
 #'
 #' @examples
-plot_gee <- function(results) {
+plot_gee_results <- function(results, yvars,
+                     xlab = "Unit difference with 95% CI in outcome for every unit increase in uVDBP and covariates") {
   results %>%
-    seer::view_main_effect(
-      groups = '~Yterms',
-      legend.title = 'P-value',
-      xlab = 'Standard deviation difference with 95% CI in the outcomes\nfor every 1 unit increase in the component',
-      ylab = 'PLS components',
-      group.label.switch = 'both'
-    ) +
-    graph_theme(ticks = FALSE)
+    dplyr::mutate(Xterms = term) %>%
+    dplyr::filter(!term == "(Intercept)") %>%
+    dplyr::mutate(Yterms = factor(Yterms,
+                                  levels = yvars,
+                                  ordered = TRUE),
+                  Xterms = factor(Xterms,
+                                  levels = rev(c("<-Xterm",
+                                                 "VN",
+                                                 "ageBase",
+                                                 "SexMale",
+                                                 "EthnicityEuropean",
+                                                 "BMI",
+                                                 "fDMdiabetes")),
+                                  labels = rev(c("uVDBP (ug/mL)",
+                                                 "Follow-up Duration (Years)",
+                                                 "Baseline Age (Years)",
+                                                 "Sex (male)",
+                                                 "Ethnicity (European",
+                                                 "BMI (kg/m^2)",
+                                                 "Diabetes")),
+                                  ordered = TRUE)) %>%
+    arrange(Xterms) %>%
+    gee_plot(xlab = xlab)
 }
-
-
-gee %>%
-  dplyr::mutate(Xterms = term) %>%
-  dplyr::filter(!term == "(Intercept)") %>%
-  dplyr::mutate(Yterms = factor(Yterms,
-                                levels = c("ACR", "eGFR"),
-                                ordered = TRUE),
-                Xterms = factor(Xterms,
-                                levels = rev(c("<-Xterm",
-                                               "VN",
-                                               "ageBase",
-                                               "SexMale",
-                                               "EthnicityEuropean",
-                                               "BMI",
-                                               "fDMdiabetes")),
-                                labels = rev(c("uVDBP (ug/mL)",
-                                               "Follow-up Duration (Years)",
-                                               "Baseline Age (Years)",
-                                               "Sex (male)",
[…]

### 2017-08-30 R/fetch_data.R
created oral contraceptive variable (0 = not taking, 1 = taking)

@@ -81,7 +81,8 @@ fetch_data <- function() {
       vitdStatus = cut(VitaminD,
                        breaks = c(-Inf, 50, 75, Inf),
                        labels = c("Deficient", "Insufficient", "Normal"),
-                       ordered = TRUE)
+                       ordered = TRUE),
+      OralContraceptive = ifelse(BirthControl == "2", 1, 0)
     ) %>%
     dplyr::filter(UDBP < 10000) %>%
     dplyr::filter(eGFR < 200) %>%
@@ -95,6 +96,7 @@ fetch_data <- function() {
       Ethnicity,
       VN,
       fVN,
+      MonthsFromBaseline,
       Glucose0,
       Glucose120,
       DM,
@@ -127,12 +129,16 @@ fetch_data <- function() {
       CRP,
       Canoe,
       BirthControl,
+      OralContraceptive,
       PeriodsStopped,
       PeriodsStoppedAge,
       MedHistHypertenAge,
       MedsBloodPressure,
       MedsDxDM,
-      MonthsFromBaseline
+      diet_milk,
+      diet_cal,
+      diet_supp_cal,
+      diet_supp_vitd
       # MedName,
       # dplyr::matches("meds")
     )

### 2017-08-31 R/progression.R
added package name to functions

@@ -42,7 +42,7 @@ plot_progress <- function(data = project_data,
                                         y = yvar)) +
     ggplot2::geom_line(data = data, ggplot2::aes_string(group = groupby), alpha = 0.7,
                        colour = "#cccccc") +
-    ggplot2::geom_line(aes(group = 1), alpha = 0.7, size = 2, colour = "#050202",
+    ggplot2::geom_line(ggplot2::aes(group = 1), alpha = 0.7, size = 2, colour = "#050202",
                        stat = "summary", fun.y = mean) +
     ggplot2::ylab(ylab)
 }

### 2017-08-31 R/gee.R
gee functions specifically for vitamin d analysis

@@ -3,6 +3,29 @@
 
 # Grab or combine data ----------------------------------------------------
 
+#' Prepare the project data for analysis through GEE. The output is a dataframe
+#' with baseline values.
+#'
+#' @param data
+#'
+#' @export
+prep_mason_data <- function(data) {
+  data %>%
+    dplyr::mutate(
+      UDBP = UDBP/1000, # udbp units to ug/mL
+      udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
+      ageBase = ifelse(fVN == "Baseline", Age, NA),
+      DM = ifelse(DM == 1, "diabetes", "non_dia"),
+      fDM = relevel(as.factor(DM), "non_dia"),
+      Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
+      Ethnicity = relevel(as.factor(Ethnicity), "Other")
+    ) %>%
+    dplyr::arrange(SID, fVN) %>%
+    dplyr::group_by(SID) %>%
+    tidyr::fill(udbpBase, ageBase) %>%
+    dplyr::ungroup() %>%
+    dplyr::arrange(SID, VN)
+}
 
 #' Prepare the project data for analysis through GEE. The output is a dataframe
 #' with baseline values. Subjects with macroalbuminuria and moderate kidney
@@ -10,7 +33,7 @@
 #'
 #' @param data project data
 #' @export
-prep_mason_data <- function(data) {
+prep_mason_data_kidney <- function(data) {
   data %>%
     dplyr::mutate(
       UDBP = UDBP/1000, # udbp units to ug/mL
@@ -124,7 +147,7 @@ explore_gee <- function(data = project_data,
 #' @export
 #'
 #' @examples
-plot_gee_results <- function(results, yvars,
+plot_gee_results_kidney <- function(results, yvars,
                      xlab = "Unit difference with 95% CI in outcome for every unit increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
@@ -140,12 +163,56 @@ plot_gee_results <- function(results, yvars,
                                                  "EthnicityEuropean",
                                                  "BMI",
                                                  "fDMdiabetes")),
-                                  labels = rev(c("uVDBP (ug/mL)",
+                                  labels = rev(c("Baseline uVDBP (ug/mL)",
                                                  "Follow-up Duration (Years)",
                                                  "Baseline Age (Years)",
                                                  "Sex (male)",
[…]

### 2017-09-14 R/gee.R
removed subjects with diabets at baseline; added PreDM variable

@@ -41,12 +41,15 @@ prep_mason_data_kidney <- function(data) {
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "diabetes", "non_dia"),
       fDM = relevel(as.factor(DM), "non_dia"),
+      PreDM = ifelse(dmStatus == "PreDM", "PreDM", "nPreDm"),
+      fPreDM = relevel(as.factor(PreDM), "nPreDM"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other")
     ) %>%
     dplyr::filter(!(fVN == "Baseline" &
                       acrStatus == "Macroalbuminuria")) %>%
     dplyr::filter(!(fVN == "Baseline" & eGFRStatus == "Moderate")) %>%
+    dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
     tidyr::fill(udbpBase, ageBase) %>%

### 2017-09-14 R/gee.R
gee plot corresponding to new gee analysis

@@ -41,8 +41,8 @@ prep_mason_data_kidney <- function(data) {
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "diabetes", "non_dia"),
       fDM = relevel(as.factor(DM), "non_dia"),
-      PreDM = ifelse(dmStatus == "PreDM", "PreDM", "nPreDm"),
-      fPreDM = relevel(as.factor(PreDM), "nPreDM"),
+      fPreDM = ifelse(dmStatus == "PreDM", "PreDM", "nPreDm"),
+      fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "nDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other")
     ) %>%
@@ -164,15 +164,15 @@ plot_gee_results_kidney <- function(results, yvars,
                                                  "ageBase",
                                                  "SexMale",
                                                  "EthnicityEuropean",
-                                                 "BMI",
-                                                 "fDMdiabetes")),
+                                                 "fPreDMPreDM",
+                                                 "fDysglycemianDysglycemia")),
                                   labels = rev(c("Baseline uVDBP (ug/mL)",
                                                  "Follow-up Duration (Years)",
                                                  "Baseline Age (Years)",
                                                  "Sex (male)",
-                                                 "Ethnicity (European",
-                                                 "BMI (kg/m^2)",
-                                                 "Diabetes")),
+                                                 "Ethnicity (European)",
+                                                 "Pre-Diabetes",
+                                                 "Dysglycemia")),
                                   ordered = TRUE)) %>%
     arrange(Xterms) %>%
     gee_plot(xlab = xlab)

### 2017-09-18 R/gee.R
reversed output on dysglycemia

@@ -41,10 +41,11 @@ prep_mason_data_kidney <- function(data) {
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "diabetes", "non_dia"),
       fDM = relevel(as.factor(DM), "non_dia"),
-      fPreDM = ifelse(dmStatus == "PreDM", "PreDM", "nPreDm"),
-      fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "nDysglycemia"),
+      fPreDM = ifelse(dmStatus == "PreDM", "PreDM", "notPreDM"),
+      fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "notDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
-      Ethnicity = relevel(as.factor(Ethnicity), "Other")
+      Ethnicity = relevel(as.factor(Ethnicity), "Other"),
+      fDysglycemia = relevel(as.factor(fDysglycemia), "notDysglycemia")
     ) %>%
     dplyr::filter(!(fVN == "Baseline" &
                       acrStatus == "Macroalbuminuria")) %>%
@@ -122,7 +123,7 @@ mason_gee <- function(data = project_data,
     } %>%
     mason::construct() %>%
     mason::scrub() %>%
-    mason::polish_renaming(rename_variables)
+    mason::polish_renaming(rename_gee_kidney)
   # %>%
   #   mason::polish_filter(extract_term, "term")
 }

### 2017-09-18 R/rename.R
separated gee renaming function from subject characteristic table renaming function

@@ -1,6 +1,4 @@
 
-# Renaming ----------------------------------------------------------------
-
 #' Renames variables in data set to improve understanding and include units
 #'
 #' @param x
@@ -22,6 +20,7 @@ rename_variables <- function(x) {
     gsub('UrineMicroalbumin', 'Urinary Microalbumin (mg/L)', .) %>%
     gsub('VitaminD', 'Serum 25(OH)D (nmol/L)', .) %>%
     gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
+    gsub("udbpBase", "Baseline uVDBP (ug/mL)", .) %>%
     gsub('udbpCrRatio', 'uVDBP:creatinine (ug/mmol)', .) %>%
     gsub('Diastolic', 'Diastolic Blood Pressure (mmHg)', .) %>%
     gsub('MeanArtPressure', 'Mean Arterial Pressure (mmHg)', .) %>%
@@ -31,8 +30,35 @@ rename_variables <- function(x) {
     gsub('Glucose0', 'Fasting', .) %>%
     gsub('Glucose120', '2h OGTT', .) %>%
     gsub('dm_status', 'Diabetic Status', .) %>%
-    gsub('DM', 'Diabetes', .) %>%
     gsub("fPreDMPreDM", "Prediabetes", .) %>%
+    gsub('DM', 'Diabetes', .) %>%
     gsub('NGT', 'Normal Glucose Tolerance', .) %>%
     gsub("fDysglycemianDysglycemia", "Dysglycemia", .)
+}
+
+
+#' Renames variables in kidney measures GEE results
+#'
+#' @param x
+#'
+#' @return
+#' @export
+#'
+#' @examples
+rename_gee_kidney <- function(x) {
+  x %>%
+    gsub("<-Xterm", "Baseline uVDBP (ug/mL)", .) %>%
+    gsub("VN", "Follow-up Duration (years)", .) %>%
+    gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
+    gsub('Age', 'Age (years)', .) %>%
+    gsub("ageBase", "Baseline Age (years)", .) %>%
+    gsub('eGFR', 'Estimated GFR (ml/min/1.73m^2)', .) %>%
+    gsub('ACR', 'Urinary albumin:creatinine (mg/mmol)', .) %>%
[…]

### 2017-09-18 R/rename.R
shortened renaming for acr and egfr

@@ -30,10 +30,8 @@ rename_variables <- function(x) {
     gsub('Glucose0', 'Fasting', .) %>%
     gsub('Glucose120', '2h OGTT', .) %>%
     gsub('dm_status', 'Diabetic Status', .) %>%
-    gsub("fPreDMPreDM", "Prediabetes", .) %>%
     gsub('DM', 'Diabetes', .) %>%
-    gsub('NGT', 'Normal Glucose Tolerance', .) %>%
-    gsub("fDysglycemianDysglycemia", "Dysglycemia", .)
+    gsub('NGT', 'Normal Glucose Tolerance', .)
 }
 
 
@@ -52,8 +50,8 @@ rename_gee_kidney <- function(x) {
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
-    gsub('eGFR', 'Estimated GFR (ml/min/1.73m^2)', .) %>%
-    gsub('ACR', 'Urinary albumin:creatinine (mg/mmol)', .) %>%
+    gsub('eGFR', 'eGFR (ml/min/1.73m^2)', .) %>%
+    gsub('ACR', 'uAlbumin:creatinine (mg/mmol)', .) %>%
     gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
     gsub("udbpBase", "Baseline uVDBP (ug/mL)", .) %>%
     gsub('udbpCrRatio', 'uVDBP:creatinine (ug/mmol)', .) %>%

### 2017-09-18 R/gee.R
removed explore_gee() from luke's code

@@ -129,20 +129,6 @@ mason_gee <- function(data = project_data,
 }
 
 
-explore_gee <- function(data = project_data,
-                        covars = covariates, caption = NULL, graph = FALSE) {
-  m <- analyze_gee(data = data, covars = covars)
-  m <- filter(m, p.value <= 0.05)
-  print(nrow(m))
-  if (graph) {
-    print(caption)
-    plot_gee_main(m)
-  } else {
-    table_gee_main(m, caption = caption)
-  }
-}
-
-
 # Plotting ----------------------------------------------------------------
 
 #' Plot GEE results in a forest plot-style
@@ -161,13 +147,13 @@ plot_gee_results_kidney <- function(results, yvars,
                                   levels = yvars,
                                   ordered = TRUE),
                   Xterms = factor(Xterms,
-                                  levels = rev(c("<-Xterm",
-                                                 "MonthsFromBaseline",
-                                                 "ageBase",
+                                  levels = rev(c("Baseline uVDBP (ug/mL)",
+                                                 "Follow-up Duration (months)",
+                                                 "Baseline Age (years)",
                                                  "SexMale",
                                                  "EthnicityEuropean",
-                                                 "fPreDMPreDM",
-                                                 "fDysglycemianDysglycemia")),
+                                                 "Prediabetes",
+                                                 "Dysglycemia")),
                                   labels = rev(c("Baseline uVDBP (ug/mL)",
                                                  "Follow-up Duration (Months)",
                                                  "Baseline Age (Years)",

### 2017-09-19 doc/vitaminD.Rmd
init commit

---
title: 'Paper 2: Vitamin D'
author: "Windy Wang"
output: word_document
---

```{r setup, collapse=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

# Only set if the Rmd file is not in the parent directory (ie. 'projectname/')
knitr::opts_knit$set(root.dir = '../')

knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
devtools::load_all()
load_data(update = TRUE)
set_options()
# extrafont::loadfonts(device="win")
```

```{r loadData, message=FALSE}

# source('.Rprofile')
# run_setup()
# load_data('data/project_data.rda')
# dim(ds)
ds <- project_data

dsBase <- ds %>% 
  dplyr::filter(fVN == "Baseline")

# Subjects with measurements at all visit numbers

dsComplete <- ds %>%
  dplyr::group_by(SID) %>%
  dplyr::filter(n() == 3) %>%
  dplyr::ungroup()

```

PAPER 2: VITAMIN D RESULTS
==========================

### Subject Characteristics

```{r subchar_vitd_baseline}

# Diet information is not available at baseline. Only VN 3 data is available at the present time.

subchar_vitd_base <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "udbpCrRatio",
           "VitaminD",
           "PTH",
           "MET",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "diet_milk",
           "diet_cal",
           "diet_supp_cal",
           "diet_supp_vitd",
           "OralContraceptive",
           "dmStatus"),
  strata = c("vitdStatus"),
  data = dsBase,
  factorVars = c("Sex", "Ethnicity", "OralContraceptive", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

```{r subchar_vitd_time}

# Diet information is not available at baseline. Only VN 3 data is available at the present time.

subchar_vitd_base <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "VitaminD",
           "PTH",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "diet_milk",
           "diet_cal",
           "diet_supp_cal",
           "diet_supp_vitd",
           "OralContraceptive",
           "dmStatus"),
  strata = c("fVN"),
  data = ds,
  factorVars = c("Sex", "Ethnicity", "OralContraceptive", "dmStatus")
) %>% 
  print(nonnormal = c("UDBP",
                      "ACR",
                      "UrineMicroalbumin"),
        quote = FALSE,
        noSpaces = TRUE)
```

### Cross-Sectional

```{r vitd_boxplot}

# Clean data
vitd <- ds %>% 
  filter(fVN == "Baseline") %>% 
  select(vitdStatus, udbpCrRatio) %>%
  na.omit()

# Box plot of uVDBP in different albuminuria categories
vitd %>% 
  box_plot("vitdStatus", "log(udbpCrRatio)", 
            "Vitamin D Status",
            "log uVDBP:Creatinine")

# n
vitd %>% 
  group_by(vitdStatus) %>% 
  summarise(n = n())

# ANOVA
anova <- aov(formula = log(udbpCrRatio)~vitdStatus, data = vitd)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r vitd_scatterplot}

# Scatterplot of ACR and uVDBP ----------------------------------

dsBase %>% 
  scatter_plot("log(udbpCrRatio)", "VitaminD", 
               "log UDBP:Creatinine",
               "Serum 25(OH)D")

# Spearman Correlation ------------------------------------------

dsBase %>% 
  cor.test(formula = ~ VitaminD + udbpCrRatio, data = ., method = "spearman")

# Linear Regression ---------------------------------------------

dsBase %>% 
  prep_mason_data() %>% 
  mason_glm(y = "VitaminD",
            x = "udbpCrRatio",
            covars = c("ageBase", "Sex", "Ethnicity", "MET", "BMI", "Season", "fDM")
            ) %>% 
  dplyr::filter(!term == "(Intercept)") %>%
  dplyr::mutate(p = round(p.value, 2),
    p = ifelse(p == "0", "<0.001", p),
    estCI = paste0(round(estimate, 2), " (",
                               round(conf.low, 2), ", ",
                               round(conf.high, 2), ")")) %>% 
  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
  # tidyr::spread(Yterms, estCI) %>%
  pander::pander()

```

PTH

```{r vitd_pth}

# Scatterplot

dsBase %>% 
  scatter_plot("PTH", "VitaminD", 
               "Parathyroid Hormone (pmol/L)",
               "Serum 25(OH)D (nmol/L)")

# Linear Regression ---------------------------------------------

dsBase %>% 
  prep_mason_data() %>% 
  mason_glm(y = "VitaminD",
            x = "PTH"
            ) %>% 
  dplyr::filter(!term == "(Intercept)") %>%
  dplyr::mutate(p = round(p.value, 2),
    p = ifelse(p == "0", "<0.001", p),
    estCI = paste0(round(estimate, 2), " (",
                               round(conf.low, 2), ", ",
                               round(conf.high, 2), ")")) %>% 
  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
  # tidyr::spread(Yterms, estCI) %>%
  pander::pander()
```

### Progression

```{r vitd_progression}

ds %>% 
  plot_progress(yvar = "VitaminD",
                ylab = "Serum 25(OH)D")

# Complete data

dsComplete %>% 
  plot_progress(yvar = "VitaminD",
                ylab = "Serum 25(OH)D")
```

### Generalized Estimating Equations

```{r vitD_gee_baseline}

# Predictor is baseline UDBP ----------------------------------------

gee_vitd_baseline <- ds %>% 
  prep_mason_data_kidney() %>% 
  mason_gee(y = c("VitaminD"),
            x = "udbpBase",
            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "dmStatus", "Season"),
            intvar = "Season")

# GEE table ----------------------------------------------------------

gee_vitd_baseline %>% 
  dplyr::filter(!term == "(Intercept)") %>% 
  dplyr::mutate(p = round(p.value, 2),
    p = ifelse(p == "0", "<0.001", p),
    estCI = paste0(round(estimate, 2), " (",
                               round(conf.low, 2), ", ",
                               round(conf.high, 2), ")")) %>% 
  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
  # tidyr::spread(Yterms, estCI) %>%
  pander::pander()

# Plot ---------------------------------------------------------------

plot_gee_results_vitd(gee_vitd_baseline,
                 yvars = c("uAlbumin:creatinine (mg/mmol)", 
                           "eGFR (ml/min/1.73m^2)"))
```

```{r vitD_gee}

# Predictor is baseline UDBP ----------------------------------------

gee <- ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("VitaminD"),
            x = "udbpBase",
            covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "PTH", "fDM"))
  # dplyr::filter(!term == "(Intercept)") %>%
  # dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error, sample.total)
  # knitr::kable()

# Predictor is UDBP at all time points ------------------------------

gee <- ds %>% 
  prep_mason_data() %>% 
  mason_gee(y = c("VitaminD"),
              x = "UDBP",
              covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "PTH", "fDM")) %>%
  dplyr::filter(!term == "(Intercept)") %>%
  dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error, sample.total)
```

```{r gee_plot}

plot_gee_results_vitd(gee,
                 yvars = c("VitaminD"))
```

### 2017-09-19 R/plots.R
plot for interactions

@@ -189,10 +189,10 @@ scatter_plot = function(data=project_data,
 
   if(line == TRUE & facet == FALSE) {
     myplot +
-      ggplot2::geom_smooth(method = loess, colour = "grey50")
+      ggplot2::geom_smooth(method = lm, colour = "grey50")
   } else if(line == TRUE & facet == TRUE) {
     myplot +
-      ggplot2::geom_smooth(method = loess, colour = "grey50") +
+      ggplot2::geom_smooth(method = lm, colour = "grey50") +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else if(line == FALSE & facet == TRUE) {
     myplot +
@@ -342,4 +342,35 @@ gee_plot = function(gee_results, xlab = "") {
                    axis.title = ggplot2::element_text(size = 10)) + #family = "Montserrat"
     ggplot2::ylab('') +
     ggplot2::xlab(xlab)
+}
+
+
+# Interaction -------------------------------------------------------------
+
+#' Plot interaction terms
+#'
+#' @param data
+#' @param xvar
+#' @param yvar
+#' @param intvar
+#' @param xlab
+#' @param ylab
+#'
+#' @return
+#'
+#' @examples
+interac_plot = function(data = project_data,
+                        xvar = "log(udbpCrRatio)",
+                        yvar,
+                        intvar,
+                        xlab = "log uVDBP:creatinine (ug/mmol)",
+                        ylab
+                        ) {
+  ggplot2::ggplot(data, ggplot2::aes_string(x = xvar,
+                                            y = yvar,
+                                            colour = intvar)) +
+  ggplot2::geom_point(colour = "grey", alpha = 0.5) +
+  ggplot2::geom_smooth(method = "lm") +
+  ggplot2::xlab(xlab) +
+  ggplot2::ylab(ylab)
 }

### 2017-09-19 doc/vitaminD.Rmd
plot interaction terms

@@
 gee_vitd_baseline <- ds %>% 
-  prep_mason_data_kidney() %>% 
+  prep_mason_data_vitd() %>% 
   mason_gee(y = c("VitaminD"),
@@
             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "dmStatus", "Season"),
-            intvar = "Season")
+            intvar = "Season") %>% 
+  mason::polish_renaming(rename_gee_vitd)
 
@@
 plot_gee_results_vitd(gee_vitd_baseline,
-                 yvars = c("uAlbumin:creatinine (mg/mmol)", 
-                           "eGFR (ml/min/1.73m^2)"))
+                 yvars = "Serum 25(OH)D (nmol/L)")
+
+# Plot interaction ---------------------------------------------------
+
+dsBase %>% 
+  ggplot2::ggplot(ggplot2::aes(x = log(udbpCrRatio), y = VitaminD, colour = Season)) +
+  ggplot2::geom_point(colour = "grey", alpha = 0.5) +
+  ggplot2::geom_smooth(method = "lm") +
+  ggplot2::xlab("log uVDBP:creatinine (ug/mmol)") +
+  ggplot2::ylab("Serum 25(OH)D (nmol/L)")
+```
+
+```{r}
+gee_vitd_baseline <- ds %>% 
+  prep_mason_data_vitd() %>% 
+  mason_gee(y = c("VitaminD"),
+            x = "udbpBase",
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "dmStatus", "PTH"),
+            intvar = "PTH") %>% 
+  mason::polish_renaming(rename_gee_vitd)
+
+# Plot interaction ---------------------------------------------------
+
+dsBase %>% 
+  interac_plot(yvar = "VitaminD",
+               intvar = "Season",
+               ylab = "Serum 25(OH)D (nmol/L)")
 ```

### 2017-09-19 R/gee.R
added MET into vitD gee plot

@@ -240,7 +240,8 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "Baseline Age (years)",
                                                  "SexMale",
                                                  "EthnicityEuropean",
-                                                 "BMI (kg/m^2) (kg/m^2)",
+                                                 "MET",
+                                                 "BMI (kg/m^2)",
                                                  "dmStatusPreDiabetes",
                                                  "dmStatusDiabetes",
                                                  "SeasonWinter",
@@ -250,6 +251,7 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "Baseline Age (years)",
                                                  "Sex (male)",
                                                  "Ethnicity (European)",
+                                                 "MET (kcal/kg/h)",
                                                  "BMI (kg/m^2)",
                                                  "Prediabetes",
                                                  "Diabetes",

### 2017-09-19 R/rename.R
renamed PTH and MET

@@ -87,5 +87,6 @@ rename_gee_vitd <- function(x) {
     gsub('DM', 'Diabetes', .) %>%
     gsub('NGT', 'Normal Glucose Tolerance', .) %>%
     gsub("fDysglycemiaDysglycemia", "Dysglycemia", .) %>%
-    gsub("PTH", "Parathyroid Hormone (pmol/L)", .)
+    gsub("PTH", "PTH (pmol/L)", .) %>%
+    gsub("MET", "MET (kcal/kg/h)", .)
 }

### 2017-09-19 doc/vitaminD.Rmd
finished vitamin D table shells analysis

--- 2017-09-19 doc/vitaminD.Rmd
@@
 
+```{r caption_setup}
+
+## Include captions below using `captioner` package
+
+fig_nums <- captioner::captioner(prefix = 'Figure')
+supfig_nums <- captioner::captioner(prefix = 'Supplementary Figure')
+tab_nums <- captioner::captioner(prefix = 'Table')
+suptab_nums <- captioner::captioner(prefix = 'Supplementary Table')
+```
+
 PAPER 2: VITAMIN D RESULTS
@@
 # Diet information is not available at baseline. Only VN 3 data is available at the present time.
+
+tab_nums("subchar_vitd_baseline", "Subject characteristics across vitamin D status at baseline")
 
@@
         quote = FALSE,
-        noSpaces = TRUE)
+        noSpaces = TRUE) %>% 
+  knitr::kable()
 ```
@@
 # Diet information is not available at baseline. Only VN 3 data is available at the present time.
+
+tab_nums("subchar_vitd", "Subject characteristics across visit numbers")
 
@@
            "Waist",
-           "eGFR",
-           "ACR",
-           "UrineCreatinine",
-           "UrineMicroalbumin",
-           "UrinaryCalcium",
-           "UDBP",
+           "MET",
            "VitaminD",
            "PTH",
+           "UDBP",
+           "udbpCrRatio",
            "Systolic",
@@
            "MeanArtPressure",
-           "diet_milk",
-           "diet_cal",
-           "diet_supp_cal",
-           "diet_supp_vitd",
            "OralContraceptive",
@@
   print(nonnormal = c("UDBP",
-                      "ACR",
-                      "UrineMicroalbumin"),
+                      "udbpCrRatio"),
         quote = FALSE,
@@
   mason_glm(y = "VitaminD",
-            x = "udbpCrRatio",
-            covars = c("ageBase", "Sex", "Ethnicity", "MET", "BMI", "Season", "fDM")
+            x = "logudbpCrRatio",
+            covars = c("ageBase", "Sex", "Ethnicity", "BMI", "MET", "fDM")
             ) %>% 
@@
 
-PTH
+#### Possible Effect Modifiers
 
@@
 
+```{r vitd_season}
+
+# Box plot of 25(OH)D in different seasons
+dsBase %>% 
+  box_plot("Season", "VitaminD", 
+            "Seasons",
+            "Serum 25(OH)D (nmol/L)")
+
+# n
+dsBase %>% 
+  group_by(Season) %>% 
+  summarise(n = n())
+
+# ANOVA
+anova <- aov(formula = VitaminD~Season, data = dsBase)
+summary(anova)
+TukeyHSD(anova)
+rm(anova)
+```
+
 ### Progression
@@
                 ylab = "Serum 25(OH)D")
+
+# ANOVA ------------------------------------------------------------
+
+anova <- aov(formula = VitaminD~fVN, data = dsComplete)
+summary(anova)
+TukeyHSD(anova)
+rm(anova)
 ```
@@
 gee_vitd_baseline <- ds %>% 
-  prep_mason_data_kidney() %>% 
-  mason_gee(y = c("VitaminD"),
+  prep_mason_data_vitd() %>% 
+  mason_gee(y = c("VitaminD", "PTH"),
             x = "udbpBase",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI", "dmStatus", "Season"),
-            intvar = "Season")
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            intvar = "Season") %>% 
+  mason::polish_renaming(rename_gee_vitd)
 
@@
 
+gee_vitd_baseline <- ds %>% 
+  prep_mason_data_vitd() %>% 
+  mason_geeplot(y = c("VitaminD", "PTH"),
+            x = "udbpBase",
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            intvar = "Season") %>% 
+  mason::polish_renaming(rename_gee_vitd)
+
 plot_gee_results_vitd(gee_vitd_baseline,
-                 yvars = c("uAlbumin:creatinine (mg/mmol)", 
-                           "eGFR (ml/min/1.73m^2)"))
+                 yvars = c("Serum 25(OH)D (nmol/L)",
+                           "PTH (pmol/L)"))
+
+# Plot interaction ---------------------------------------------------
+
+dsBase %>% 
+  interac_plot(yvar = "PTH",
+               intvar = "Season",
+               ylab = "Parathyroid Hormone (pmol/L)") #Serum 25(OH)D (nmol/L) #Parathyroid Hormone (pmol/L)
 ```
@@
 
-# Predictor is baseline UDBP ----------------------------------------
+# Predictor is UDBP ----------------------------------------
 
-gee <- ds %>% 
-  prep_mason_data() %>% 
-  mason_gee(y = c("VitaminD"),
-            x = "udbpBase",
-            covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "PTH", "fDM"))
-  # dplyr::filter(!term == "(Intercept)") %>%
-  # dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error, sample.total)
-  # knitr::kable()
+gee_vitd <- ds %>% 
+  prep_mason_data_vitd() %>% 
+  mason_gee(y = c("VitaminD", "PTH"),
+            x = "UDBP",
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            intvar = "Season") %>% 
+  mason::polish_renaming(rename_gee_vitd)
 
-# Predictor is UDBP at all time points ------------------------------
+# GEE table ----------------------------------------------------------
 
-gee <- ds %>% 
-  prep_mason_data() %>% 
-  mason_gee(y = c("VitaminD"),
+gee_vitd %>% 
+  dplyr::filter(!term == "(Intercept)") %>% 
+  dplyr::mutate(p = round(p.value, 2),
+    p = ifelse(p == "0", "<0.001", p),
+    estCI = paste0(round(estimate, 2), " (",
+                               round(conf.low, 2), ", ",
+                               round(conf.high, 2), ")")) %>% 
+  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
+  # tidyr::spread(Yterms, estCI) %>%
+  pander::pander()
+
+# Plot ---------------------------------------------------------------
+
+gee_vitd <- ds %>% 
+  prep_mason_data_vitd() %>% 
+  mason_geeplot(y = c("VitaminD", "PTH"),
             x = "UDBP",
-              covars = c("VN", "ageBase", "Sex", "Ethnicity", "BMI", "PTH", "fDM")) %>%
-  dplyr::filter(!term == "(Intercept)") %>%
-  dplyr::select(Yterms, Xterms, term, estimate, p.value, conf.low, conf.high, std.error, sample.total)
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            intvar = "Season") %>% 
+  mason::polish_renaming(rename_gee_vitd)
+
+plot_gee_results_vitd(gee_vitd,
+                 yvars = c("Serum 25(OH)D (nmol/L)",
+                           "PTH (pmol/L)"))
+
+# Plot interaction ---------------------------------------------------
+
+ds %>% 
+  interac_plot(yvar = "VitaminD",
+               intvar = "Season",
+               ylab = "Serum 25(OH)D (nmol/L)")
 ```
-
-```{r gee_plot}
-
-plot_gee_results_vitd(gee,
-                 yvars = c("VitaminD"))
-```

### 2017-09-22 R/gee.R
scaled vitamin d covariates

@@ -69,9 +69,10 @@ prep_mason_data_kidney <- function(data) {
 #'
 #' @export
 prep_mason_data_vitd <- function(data) {
+
   data %>%
     dplyr::mutate(
-      UDBP = UDBP/1000, # udbp units to ug/mL
+      # UDBP = UDBP/1000, # udbp units to ug/mL
       udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "DM", "notDM"),
@@ -82,11 +83,20 @@ prep_mason_data_vitd <- function(data) {
     ) %>%
     dplyr::filter(!(fVN == "Baseline" & vitdStatus == "Deficient")) %>%
     dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
+
+    dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
+                       UDBP,
+                       udbpBase,
+                       MonthsFromBaseline,
+                       ageBase,
+                       MET,
+                       BMI) %>%
+
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
     tidyr::fill(udbpBase, ageBase) %>%
     dplyr::ungroup() %>%
-    dplyr::arrange(SID, VN)
+    dplyr::arrange(SID, fVN)
 }
 
 # Analyze -----------------------------------------------------------------
@@ -134,9 +144,21 @@ mason_gee <- function(data = project_data,
       }
     } %>%
     mason::scrub()
+    # mason::polish_transform_estimates(function(x) (exp(x) - 1) * 100)
 }
 
 
+#' Run GEE on prepared project data simplified for graphical purposes (plot_gee_results)
+#'
+#' @param data
+#' @param y
+#' @param x
+#' @param covars
+#' @param intvar
+#'
+#' @return
+#'
+#' @examples
 mason_geeplot <- function(data = project_data,
                       y = outcomes,
                       x = predictors,
@@ -172,6 +194,7 @@ mason_geeplot <- function(data = project_data,
     } %>%
     mason::construct() %>%
     mason::scrub()
+    # mason::polish_transform_estimates(function(x) (exp(x) - 1) * 100)
 }
 
 
@@ -227,7 +250,7 @@ plot_gee_results_kidney <- function(results, yvars,
 #'
 #' @examples
 plot_gee_results_vitd <- function(results, yvars,
[…]

### 2017-09-22 R/plots.R
removed data specification from interac_plot

@@ -359,7 +359,7 @@ gee_plot = function(gee_results, xlab = "") {
 #' @return
 #'
 #' @examples
-interac_plot = function(data = project_data,
+interac_plot = function(data,
                         xvar = "log(udbpCrRatio)",
                         yvar,
                         intvar,

### 2017-09-22 R/regression.R
added interaction value to regression

@@ -12,9 +12,17 @@
 mason_glm <- function(data = project_data,
                         y = outcomes,
                         x = predictors,
-                        covars = NULL) {
+                        covars = NULL,
+                        intvar = NULL) {
 
   co <- !is.null(covars)
+
+  int <- !is.null(intvar)
+  if (int) {
+    extract_term <- ":"
+  } else {
+    extract_term <- "Xterm$"
+  }
 
   data %>%
     mason::design("glm") %>%
@@ -23,11 +31,17 @@ mason_glm <- function(data = project_data,
     mason::add_variables("xvars", x) %>%
     mason::construct() %>% {
       if (co) {
-        mason::add_variables(., "covariates", covars) %>%
-        mason::construct()
+        mason::add_variables(., "covariates", covars) %>% {
+          if (int) {
+            mason::add_variables(., "interaction", intvar)
+          } else {
+            .
+          }
+        }
       } else {
         .
       }
     } %>%
+    mason::construct() %>%
     mason::scrub()
 }

### 2017-09-22 doc/vitaminD.Rmd
added functions to figure out values for graphs

@@
         quote = FALSE,
-        noSpaces = TRUE)
+        noSpaces = TRUE) %>% 
+  knitr::kable()
+
+ds %>%
+  dplyr::select(VN, PTH) %>%
+  dplyr::group_by(VN) %>%
+  na.omit() %>%
+  summarise(n = n())
 ```
@@
             "Vitamin D Status",
-            "log uVDBP:Creatinine")
+            "log uVDBP:Creatinine (μg/mmol)")
+
+# value
+vitd %>%
+  dplyr::group_by(vitdStatus) %>% 
+  dplyr::summarise(Median = median(log(udbpCrRatio)), 
+                   IQR = IQR(log(udbpCrRatio))
+                   )
 
@@
 
+tab_nums(gee_vitd_baseline, "GEE results for uVDBP at baseline (predictor) with 25(OH)D and (outcomes), adjusted for follow-up duration (in months), baseline age, sex, ethnicity, physical activity (MET), BMI, diabetes status, and seasonality; subjects with deficient vitamin D status at baseline excluded (n = 713).")
+
 gee_vitd_baseline %>% 
@@
                ylab = "Parathyroid Hormone (pmol/L)") #Serum 25(OH)D (nmol/L) #Parathyroid Hormone (pmol/L)
+
+rms::ols(VitaminD ~ udbpCrRatio * Season, data = dsBase)
 ```
@@
                ylab = "Serum 25(OH)D (nmol/L)")
+
+rms::ols(VitaminD ~ udbpCrRatio * Season, data = ds)
 ```

### 2017-09-23 R/gee.R
function to display GEE results in a nice table

@@ -175,6 +175,26 @@ mason_geeplot <- function(data = project_data,
 }
 
 
+# Table -------------------------------------------------------------------
+
+#' Display GEE results in a table. Use pander::pander() to print as a table when knitting document.
+#'
+#' @param results
+#'
+#' @return
+#'
+#' @examples
+gee_results_table <- function(results) {
+  results %>%
+    dplyr::filter(!term == "(Intercept)") %>%
+    dplyr::mutate(p = round(p.value, 2),
+                  p = ifelse(p == "0", "<0.001", p),
+                  estCI = paste0(round(estimate, 2), " (",
+                                 round(conf.low, 2), ", ",
+                                 round(conf.high, 2), ")")) %>%
+    dplyr::select(Yterms, Xterms, term, estCI, p)
+}
+
 # Plotting ----------------------------------------------------------------
 
 #' Plot GEE results in a forest plot-style

### 2017-09-23 R/plots.R
removed font specifications in basic plots

@@ -179,7 +179,7 @@ scatter_plot = function(data=project_data,
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    panel.grid.minor.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 11, family = "Open Sans"), #family = "Open Sans"
+                   text = ggplot2::element_text(size = 11), #family = "Open Sans"
                    axis.text.y = ggplot2::element_text(colour = "grey30"), #"grey" normally
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "grey30")) + #angle = 45

### 2017-09-23 R/gee.R
removed dmStatus ordering in mason_data_prep

@@ -19,7 +19,8 @@ prep_mason_data <- function(data) {
       fDM = relevel(as.factor(DM), "notDM"),
       fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "notDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
-      Ethnicity = relevel(as.factor(Ethnicity), "Other")
+      Ethnicity = relevel(as.factor(Ethnicity), "Other"),
+      dmStatus = factor(dmStatus, ordered = FALSE)
     ) %>%
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
@@ -46,7 +47,8 @@ prep_mason_data_kidney <- function(data) {
       fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "notDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other"),
-      fDysglycemia = relevel(as.factor(fDysglycemia), "notDysglycemia")
+      fDysglycemia = relevel(as.factor(fDysglycemia), "notDysglycemia"),
+      dmStatus = factor(dmStatus, ordered = FALSE)
     ) %>%
     dplyr::filter(!(fVN == "Baseline" &
                       acrStatus == "Macroalbuminuria")) %>%

### 2017-09-23 R/plots.R
removed redundancy in if statements

@@ -47,7 +47,7 @@ box_plot <- function(data=project_data, xvar, yvar, xlab="", ylab="", facet = FA
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
-  if(facet == TRUE) {
+  if(facet) {
     myboxplot +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else {
@@ -95,7 +95,7 @@ box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
-  if(facet == TRUE) {
+  if(facet) {
     myboxplot +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else {
@@ -144,7 +144,7 @@ box_plot_poster <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
-  if(facet == TRUE) {
+  if(facet) {
     myboxplot +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else {

### 2017-09-23 R/gee.R
option to print as a table for gee_results_table

@@ -186,7 +186,7 @@ mason_geeplot <- function(data = project_data,
 #' @return
 #'
 #' @examples
-gee_results_table <- function(results) {
+gee_results_table <- function(results, table = TRUE) {
   results %>%
     dplyr::filter(!term == "(Intercept)") %>%
     dplyr::mutate(p = round(p.value, 2),
@@ -194,7 +194,13 @@ gee_results_table <- function(results) {
                   estCI = paste0(round(estimate, 2), " (",
                                  round(conf.low, 2), ", ",
                                  round(conf.high, 2), ")")) %>%
-    dplyr::select(Yterms, Xterms, term, estCI, p)
+    dplyr::select(Yterms, Xterms, term, estCI, p) %>% {
+      if(table) {
+        pander::pander()
+      } else {
+        .
+      }
+    }
 }
 
 # Plotting ----------------------------------------------------------------

### 2017-09-23 R/progression.R
roxygen documentation of plot_progress

@@ -30,8 +30,17 @@ plot_progress_by <- function(data,
 }
 
 
-
-
+#' Plot progression of UDBP over time (visit number) with no grouping variable
+#'
+#' @param data
+#' @param xvar
+#' @param yvar
+#' @param ylab
+#' @param groupby
+#'
+#' @return
+#'
+#' @examples
 plot_progress <- function(data = project_data,
                                xvar = "fVN", yvar,
                                ylab,

### 2017-09-23 R/progression.R
specified xlab, simplified ylab to be yvar unless otherwise specificied

@@ -12,7 +12,7 @@
 #' @examples
 plot_progress_by <- function(data,
                                xvar = "fVN", yvar = "log(UDBP)",
-                               ylab = "log(UDBP)",
+                               ylab = yvar,
                                byvar, groupby = "SID") {
   # m <- data %>%
   #   dplyr::group_by_(byvar, xvar) %>%
@@ -26,6 +26,7 @@ plot_progress_by <- function(data,
     ggplot2::geom_line(ggplot2::aes_string(group = byvar),
                        alpha = 0.7, size = 2,
                        stat = "summary", fun.y = mean) +
+    ggplot2::xlab("Follow-up Duration (years)") +
     ggplot2::ylab(ylab)
 }
 
@@ -43,7 +44,7 @@ plot_progress_by <- function(data,
 #' @examples
 plot_progress <- function(data = project_data,
                                xvar = "fVN", yvar,
-                               ylab,
+                               ylab = yvar,
                                groupby = "SID") {
 
   data %>%
@@ -53,6 +54,7 @@ plot_progress <- function(data = project_data,
                        colour = "#cccccc") +
     ggplot2::geom_line(ggplot2::aes(group = 1), alpha = 0.7, size = 2, colour = "#050202",
                        stat = "summary", fun.y = mean) +
+    ggplot2::xlab("Follow-up Duration (years)") +
     ggplot2::ylab(ylab)
 }
 
@@ -61,7 +63,7 @@ plot_progress <- function(data = project_data,
 
 plot_progress_boxplot <- function (data,
                                    xvar = "fVN",
-                                   yvar, ylab,
+                                   yvar, ylab = yvar,
                                    groupby = "SID") {
 
   data %>%
@@ -86,5 +88,6 @@ plot_progress_boxplot <- function (data,
                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
                   axis.text.x = ggplot2::element_text(colour = "grey30"), #angle = 45
                   axis.title = ggplot2::element_text(size = 10)) +
+  ggplot2::xlab("Follow-up Duration (years)") +
   ggplot2::ylab(ylab)
 }

### 2017-09-23 R/rename.R
shortened ACR renaming

@@ -15,7 +15,7 @@ rename_variables <- function(x) {
     gsub("BMI", "BMI (kg/m^2)", .) %>%
     gsub('Waist', 'Waist Circumference (cm)', .) %>%
     gsub('eGFR', 'Estimated GFR (ml/min/1.73m^2)', .) %>%
-    gsub('ACR', 'Urinary albumin:creatinine (mg/mmol)', .) %>%
+    gsub('ACR', 'ACR (mg/mmol)', .) %>%
     gsub('MicroalbCreatRatio', 'Microalbumin:Creatinine', .) %>%
     gsub('UrineCreatinine', 'Urinary Creatinine (mmol/L)', .) %>%
     gsub('UrineMicroalbumin', 'Urinary Microalbumin (mg/L)', .) %>%
@@ -52,7 +52,7 @@ rename_gee_kidney <- function(x) {
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
     gsub('eGFR', 'eGFR (ml/min/1.73m^2)', .) %>%
-    gsub('ACR', 'uAlbumin:creatinine (mg/mmol)', .) %>%
+    gsub('ACR', 'ACR (mg/mmol)', .) %>%
     gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
     gsub("udbpBase", "Baseline uVDBP (ug/mL)", .) %>%
     gsub('udbpCrRatio', 'uVDBP:creatinine (ug/mmol)', .) %>%

### 2017-09-23 R/gee.R
fixed gee kidney graph labels

@@ -194,13 +194,7 @@ gee_results_table <- function(results, table = TRUE) {
                   estCI = paste0(round(estimate, 2), " (",
                                  round(conf.low, 2), ", ",
                                  round(conf.high, 2), ")")) %>%
-    dplyr::select(Yterms, Xterms, term, estCI, p) %>% {
-      if(table) {
-        pander::pander()
-      } else {
-        .
-      }
-    }
+    dplyr::select(Yterms, Xterms, term, estCI, p)
 }
 
 # Plotting ----------------------------------------------------------------
@@ -226,15 +220,15 @@ plot_gee_results_kidney <- function(results, yvars,
                                                  "Baseline Age (years)",
                                                  "SexMale",
                                                  "EthnicityEuropean",
-                                                 "Prediabetes",
-                                                 "Dysglycemia")),
+                                                 "dmStatusPreDiabetes",
+                                                 "dmStatusDiabetes")),
                                   labels = rev(c("Baseline uVDBP (ug/mL)",
                                                  "Follow-up Duration (Months)",
                                                  "Baseline Age (Years)",
                                                  "Sex (male)",
                                                  "Ethnicity (European)",
                                                  "Prediabetes",
-                                                 "Dysglycemia")),
+                                                 "Diabetes")),
                                   ordered = TRUE)) %>%
     arrange(Xterms) %>%
     gee_plot(xlab = xlab)

### 2017-09-23 doc/vitaminD.Rmd
replaced gee table code with function

--- 2017-09-19 doc/vitaminD.Rmd
@@
 gee_vitd_baseline %>% 
-  dplyr::filter(!term == "(Intercept)") %>% 
-  dplyr::mutate(p = round(p.value, 2),
-    p = ifelse(p == "0", "<0.001", p),
-    estCI = paste0(round(estimate, 2), " (",
-                               round(conf.low, 2), ", ",
-                               round(conf.high, 2), ")")) %>% 
-  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
+  gee_results_table() %>% 
   # tidyr::spread(Yterms, estCI) %>%
@@
             x = "udbpBase",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", 
+                       "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
             intvar = "Season") %>% 

### 2017-10-12 R/fetch_data.R
added udbpTertile

@@ -75,6 +75,7 @@ fetch_data <- function() {
                        breaks = c(0, 1.23, 60, Inf),
                        labels = c("Trace", "Normal", "High"),
                        ordered_result = TRUE),
+      udbpTertile = ntile(UDBP, 3),
       udbpCr = UDBP / UrineCreatinine,
       logudbpCr = log(udbpCr),
       udbpCrTertile = ntile(udbpCr, 3),

### 2017-10-12 doc/vitaminD.Rmd
fixed gee scaling (percent change)

@@
            "UDBP",
-           "udbpCrRatio",
+           "udbpCr",
            "VitaminD",
@@
            "diet_supp_vitd",
-           "OralContraceptive",
            "dmStatus"),
@@
   data = dsBase,
-  factorVars = c("Sex", "Ethnicity", "OralContraceptive", "dmStatus")
+  factorVars = c("Sex", "Ethnicity", "dmStatus")
 ) %>% 
@@
 
+# for sensitivity analysis
+m <- ds %>% 
+  dplyr::filter(fVN != "6Year")
+
 subchar_vitd_base <- tableone::CreateTableOne(
@@
            "UDBP",
-           "udbpCrRatio",
+           "udbpCr",
            "Systolic",
@@
            "MeanArtPressure",
-           "OralContraceptive",
            "dmStatus"),
-  strata = c("fVN"),
-  data = ds,
-  factorVars = c("Sex", "Ethnicity", "OralContraceptive", "dmStatus")
+  strata = c("VN"),
+  data = m,
+  factorVars = c("Sex", "Ethnicity", "dmStatus")
 ) %>% 
   print(nonnormal = c("UDBP",
-                      "udbpCrRatio"),
+                      "udbpCr"),
         quote = FALSE,
@@
   filter(fVN == "Baseline") %>% 
-  select(vitdStatus, udbpCrRatio) %>%
+  select(vitdStatus, udbpCr) %>%
   na.omit()
@@
 vitd %>% 
-  box_plot("vitdStatus", "log(udbpCrRatio)", 
+  box_plot("vitdStatus", "log(udbpCr)", 
             "Vitamin D Status",
@@
   dplyr::group_by(vitdStatus) %>% 
-  dplyr::summarise(Median = median(log(udbpCrRatio)), 
-                   IQR = IQR(log(udbpCrRatio))
+  dplyr::summarise(Median = median(log(udbpCr)), 
+                   IQR = IQR(log(udbpCr))
                    )
@@
 # ANOVA
-anova <- aov(formula = log(udbpCrRatio)~vitdStatus, data = vitd)
+anova <- aov(formula = log(udbpCr)~vitdStatus, data = vitd)
 summary(anova)
@@
 dsBase %>% 
-  scatter_plot("log(udbpCrRatio)", "VitaminD", 
+  scatter_plot("log(udbpCr)", "VitaminD", 
                "log UDBP:Creatinine",
@@
 dsBase %>% 
-  cor.test(formula = ~ VitaminD + udbpCrRatio, data = ., method = "spearman")
+  cor.test(formula = ~ VitaminD + udbpCr, data = ., method = "spearman")
 
@@
   mason_glm(y = "VitaminD",
-            x = "logudbpCrRatio",
+            x = "logudbpCr",
             covars = c("ageBase", "Sex", "Ethnicity", "BMI", "MET", "fDM")
@@
 
+# gee_vitd_baseline <- ds %>% 
+#   prep_mason_data_vitd() %>% 
+#   mason_gee(y = c("VitaminD", "PTH"),
+#             x = "udbpCrBase",
+#             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+#             intvar = "Season") %>% 
+#   mason::polish_renaming(rename_gee_vitd)
+
 gee_vitd_baseline <- ds %>% 
   prep_mason_data_vitd() %>% 
-  mason_gee(y = c("VitaminD", "PTH"),
-            x = "udbpBase",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+  mason_geeplot(y = c("lVitD", "lPTH"),
+            x = "udbpCrBase",
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", 
+                       "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
             intvar = "Season") %>% 
@@
 
-gee_vitd_baseline <- ds %>% 
-  prep_mason_data_vitd() %>% 
-  mason_geeplot(y = c("VitaminD", "PTH"),
-            x = "udbpBase",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", 
-                       "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
-            intvar = "Season") %>% 
-  mason::polish_renaming(rename_gee_vitd)
-
 plot_gee_results_vitd(gee_vitd_baseline,
@@
 
-dsBase %>% 
-  interac_plot(yvar = "PTH",
-               intvar = "Season",
-               ylab = "Parathyroid Hormone (pmol/L)") #Serum 25(OH)D (nmol/L) #Parathyroid Hormone (pmol/L)
-
-rms::ols(VitaminD ~ udbpCrRatio * Season, data = dsBase)
+# dsBase %>% 
+#   interac_plot(yvar = "PTH",
+#                intvar = "Season",
+#                ylab = "Parathyroid Hormone (pmol/L)") #Serum 25(OH)D (nmol/L) #Parathyroid Hormone (pmol/L)
+# 
+# rms::ols(VitaminD ~ udbpCr * Season, data = dsBase)
 ```
@@
 
+# gee_vitd <- ds %>% 
+#   prep_mason_data_vitd() %>% 
+#   mason_gee(y = c("lVitD", "lPTH"),
+#             x = "udbpCr",
+#             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+#             intvar = "Season") %>% 
+#   mason::polish_renaming(rename_gee_vitd)
+
 gee_vitd <- ds %>% 
   prep_mason_data_vitd() %>% 
-  mason_gee(y = c("VitaminD", "PTH"),
-            x = "UDBP",
+  mason_geeplot(y = c("lVitD", "lPTH"),
+            x = "udbpCr",
             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
@@
 gee_vitd %>% 
-  dplyr::filter(!term == "(Intercept)") %>% 
-  dplyr::mutate(p = round(p.value, 2),
-    p = ifelse(p == "0", "<0.001", p),
-    estCI = paste0(round(estimate, 2), " (",
-                               round(conf.low, 2), ", ",
-                               round(conf.high, 2), ")")) %>% 
-  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
+  gee_results_table() %>% 
   # tidyr::spread(Yterms, estCI) %>%
@@
 # Plot ---------------------------------------------------------------
-
-gee_vitd <- ds %>% 
-  prep_mason_data_vitd() %>% 
-  mason_geeplot(y = c("VitaminD", "PTH"),
-            x = "UDBP",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
-            intvar = "Season") %>% 
-  mason::polish_renaming(rename_gee_vitd)
 
@@
 
-ds %>% 
-  interac_plot(yvar = "VitaminD",
-               intvar = "Season",
-               ylab = "Serum 25(OH)D (nmol/L)")
-
-rms::ols(VitaminD ~ udbpCrRatio * Season, data = ds)
+# ds %>% 
+#   interac_plot(yvar = "VitaminD",
+#                intvar = "Season",
+#                ylab = "Serum 25(OH)D (nmol/L)")
+# 
+# rms::ols(VitaminD ~ udbpCr * Season, data = ds)
 ```

### 2017-10-12 R/rename.R
adjusted for logged variables

@@ -51,8 +51,8 @@ rename_gee_kidney <- function(x) {
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
-    gsub('eGFR', 'eGFR (ml/min/1.73m^2)', .) %>%
-    gsub('ACR', 'ACR (mg/mmol)', .) %>%
+    gsub('leGFR', 'eGFR (ml/min/1.73m^2)', .) %>%
+    gsub('lACR', 'ACR (mg/mmol)', .) %>%
     gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
     gsub("udbpCrBase", "Baseline uVDBP:cr (ug/mmol)", .) %>%
     gsub('udbpCrRatio', 'uVDBP:creatinine (ug/mmol)', .) %>%
@@ -87,6 +87,7 @@ rename_gee_vitd <- function(x) {
     gsub('DM', 'Diabetes', .) %>%
     gsub('NGT', 'Normal Glucose Tolerance', .) %>%
     gsub("fDysglycemiaDysglycemia", "Dysglycemia", .) %>%
-    gsub("PTH", "PTH (pmol/L)", .) %>%
-    gsub("MET", "MET (kcal/kg/h)", .)
+    gsub("MET", "MET (kcal/kg/h)", .) %>%
+    gsub("lVitD", "Serum 25(OH)D (nmol/L)", .) %>%
+    gsub("lPTH", "PTH (pmol/L)", .)
 }

### 2017-10-12 R/gee.R
scaled outcomes for better results interpretation

@@ -39,18 +39,16 @@ prep_mason_data <- function(data) {
 prep_mason_data_kidney <- function(data) {
   data %>%
     dplyr::mutate(
-      UDBP = UDBP/1000, # udbp units to ug/mL
       udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
-
       ageBase = ifelse(fVN == "Baseline", Age, NA),
-      # DM = ifelse(DM == 1, "diabetes", "non_dia"),
-      fDM = relevel(as.factor(DM), "non_dia"),
+      DM = ifelse(DM == 1, "DM", "notDM"),
+      fDM = relevel(as.factor(DM), "notDM"),
       dmStatus = factor(dmStatus, ordered = FALSE),
-      # fPreDM = ifelse(dmStatus == "PreDM", "PreDM", "notPreDM"),
-      # fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "notDysglycemia"),
-      # fDysglycemia = relevel(as.factor(fDysglycemia), "notDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
-      Ethnicity = relevel(as.factor(Ethnicity), "Other")
+      Ethnicity = relevel(as.factor(Ethnicity), "Other"),
+
+      lACR = log(ACR),
+      leGFR = log(eGFR)
     ) %>%
 
     dplyr::filter(!(fVN == "Baseline" &
@@ -58,13 +56,15 @@ prep_mason_data_kidney <- function(data) {
     dplyr::filter(!(fVN == "Baseline" & eGFRStatus == "Moderate")) %>%
     dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
+    dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
+                       udbpCr,
+                       udbpCrBase,
+                       MonthsFromBaseline,
+                       ageBase) %>%
+
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
     tidyr::fill(udbpCrBase, ageBase) %>%
-    # dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
-    #                    udbpCrBase,
-    #                    udbpCr,
-    #                    ageBase) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(SID, fVN)
 }
[…]

### 2017-10-16 R/gee.R
sensitivity analysis using uVDBP as a predictor in case denominator creatinine influenced results

@@ -39,6 +39,7 @@ prep_mason_data <- function(data) {
 prep_mason_data_kidney <- function(data) {
   data %>%
     dplyr::mutate(
+      udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
       udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "DM", "notDM"),
@@ -57,6 +58,8 @@ prep_mason_data_kidney <- function(data) {
     dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
+                       UDBP,
+                       udbpBase,
                        udbpCr,
                        udbpCrBase,
                        MonthsFromBaseline,
@@ -64,7 +67,7 @@ prep_mason_data_kidney <- function(data) {
 
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
-    tidyr::fill(udbpCrBase, ageBase) %>%
+    tidyr::fill(udbpBase, ageBase) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(SID, fVN)
 }
@@ -83,6 +86,7 @@ prep_mason_data_vitd <- function(data) {
   data %>%
     dplyr::mutate(
       # UDBP = UDBP/1000, # udbp units to ug/mL
+      udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
       udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
       ageBase = ifelse(fVN == "Baseline", Age, NA),
       DM = ifelse(DM == 1, "DM", "notDM"),
@@ -99,6 +103,8 @@ prep_mason_data_vitd <- function(data) {
     dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
+                       UDBP,
+                       udbpBase,
                        udbpCr,
                        udbpCrBase,
                        MonthsFromBaseline,
@@ -108,7 +114,7 @@ prep_mason_data_vitd <- function(data) {
 
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
[…]

### 2017-10-16 R/rename.R

@@ -46,7 +46,7 @@ rename_variables <- function(x) {
 #' @examples
 rename_gee_kidney <- function(x) {
   x %>%
-    gsub("<-Xterm", "uVDBP:cr (ug/mmol)", .) %>%
+    gsub("<-Xterm", "Baseline uVDBP (ug/mmol)", .) %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
@@ -55,7 +55,7 @@ rename_gee_kidney <- function(x) {
     gsub('lACR', 'ACR (mg/mmol)', .) %>%
     gsub('UDBP', 'Urinary VDBP (ng/mL)', .) %>%
     gsub("udbpCrBase", "Baseline uVDBP:cr (ug/mmol)", .) %>%
-    gsub('udbpCrRatio', 'uVDBP:creatinine (ug/mmol)', .) %>%
+    gsub('udbpCrRatio', 'uVDBP:cr (ug/mmol)', .) %>%
     gsub("fPreDMPreDM", "Prediabetes", .) %>%
     gsub('DM', 'Diabetes', .) %>%
     gsub('NGT', 'Normal Glucose Tolerance', .) %>%
@@ -73,7 +73,7 @@ rename_gee_kidney <- function(x) {
 #' @examples
 rename_gee_vitd <- function(x) {
   x %>%
-    gsub("<-Xterm", "uVDBP:cr (ug/mmol)", .) %>%
+    gsub("<-Xterm", "uVDBP (ug/mmol)", .) %>%
     gsub("VitaminD", "Serum 25(OH)D (nmol/L)", .) %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%

### 2017-11-02 doc/vitaminD.Rmd
remove PTH from gee model

@@
   prep_mason_data_vitd() %>% 
-  mason_geeplot(y = c("lVitD", "lPTH"),
+  mason_geeplot(y = "lVitD",
             x = "udbpCrBase",

### 2017-11-02 R/gee.R

@@ -95,8 +95,8 @@ prep_mason_data_vitd <- function(data) {
       Ethnicity = relevel(as.factor(Ethnicity), "Other"),
       dmStatus = factor(dmStatus, ordered = FALSE),
 
-      lVitD = log(VitaminD),
-      lPTH = log(PTH)
+      lVitD = log(VitaminD)
+      # lPTH = log(PTH)
     ) %>%
 
     # dplyr::filter(!(fVN == "Baseline" & vitdStatus == "Deficient")) %>%

### 2017-11-06 doc/vitaminD.Rmd
removed PTH from vitamin D model

@@
 
-# gee_vitd_baseline <- ds %>% 
-#   prep_mason_data_vitd() %>% 
-#   mason_gee(y = c("VitaminD", "PTH"),
-#             x = "udbpCrBase",
-#             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
-#             intvar = "Season") %>% 
-#   mason::polish_renaming(rename_gee_vitd)
-
 gee_vitd_baseline <- ds %>% 
@@
 plot_gee_results_vitd(gee_vitd_baseline,
-                 yvars = c("Serum 25(OH)D (nmol/L)",
-                           "PTH (pmol/L)"))
+                 yvars = "Serum 25(OH)D (nmol/L)")
 
@@
 
-# gee_vitd <- ds %>% 
-#   prep_mason_data_vitd() %>% 
-#   mason_gee(y = c("lVitD", "lPTH"),
-#             x = "udbpCr",
-#             covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
-#             intvar = "Season") %>% 
-#   mason::polish_renaming(rename_gee_vitd)
-
 gee_vitd <- ds %>% 
   prep_mason_data_vitd() %>% 
-  mason_geeplot(y = c("lVitD", "lPTH"),
+  mason_geeplot(y = "lVitD",
             x = "udbpCr",

### 2017-11-06 R/gee.R

@@ -67,7 +67,7 @@ prep_mason_data_kidney <- function(data) {
 
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
-    tidyr::fill(udbpBase, ageBase) %>%
+    tidyr::fill(udbpBase, udbpCrBase, ageBase) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(SID, fVN)
 }
@@ -114,7 +114,7 @@ prep_mason_data_vitd <- function(data) {
 
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
-    tidyr::fill(udbpBase, ageBase) %>%
+    tidyr::fill(udbpBase, udbpCrBase, ageBase) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(SID, fVN)
 }
@@ -304,7 +304,7 @@ plot_gee_results_vitd <- function(results, yvars,
                                   levels = yvars,
                                   ordered = TRUE),
                   Xterms = factor(Xterms,
-                                  levels = rev(c("uVDBP:cr (ug/mmol)",
+                                  levels = rev(c("<-Xterm",
                                                  "Follow-up Duration (months)",
                                                  "Baseline Age (years)",
                                                  "SexMale",
@@ -314,7 +314,7 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "dmStatusPreDiabetes",
                                                  "dmStatusDiabetes",
                                                  "SeasonWinter",
-                                                 "uVDBP:cr (ug/mmol):SeasonWinter")),
+                                                 "<-Xterm:SeasonWinter")),
                                   labels = rev(c("uVDBP:cr (ug/mL)",
                                                  "Follow-up Duration (months)",
                                                  "Baseline Age (years)",
@@ -325,7 +325,7 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "Prediabetes",
                                                  "Diabetes",
                                                  "Seasonality (winter)",
-                                                 "Season:Baseline uVDBP:cr interaction")),
+                                                 "Season:uVDBP:cr interaction")),
                                   ordered = TRUE)) %>%
     arrange(Xterms) %>%
     gee_plot(xlab = xlab)

### 2017-11-06 R/rename.R

@@ -73,7 +73,6 @@ rename_gee_kidney <- function(x) {
 #' @examples
 rename_gee_vitd <- function(x) {
   x %>%
-    gsub("<-Xterm", "uVDBP (ug/mmol)", .) %>%
     gsub("VitaminD", "Serum 25(OH)D (nmol/L)", .) %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%

### 2017-11-06 R/gee.R
fixed udbpCrBase error by adding it to dplyr::fill

@@ -252,14 +252,14 @@ gee_results_table <- function(results, table = TRUE) {
 plot_gee_results_kidney <- function(results, yvars,
                      xlab = "Percent difference with 95% CI in the outcomes
                      for each SD increase in uVDBP and covariates",
-                     terms = c("Baseline uVDBP (ug/mmol)",
+                     terms = c("<-Xterm",
                                "Follow-up Duration (months)",
                                "Baseline Age (years)",
                                "SexMale",
                                "EthnicityEuropean",
                                "dmStatusPreDiabetes",
                                "dmStatusDiabetes"),
-                     labels = c("Baseline uVDBP (ug/mL)",
+                     labels = c("uVDBP:cr (ug/mL)",
                                 "Follow-up Duration (Months)",
                                 "Baseline Age (Years)",
                                 "Sex (male)",

### 2017-11-06 R/rename.R

@@ -46,7 +46,6 @@ rename_variables <- function(x) {
 #' @examples
 rename_gee_kidney <- function(x) {
   x %>%
-    gsub("<-Xterm", "Baseline uVDBP (ug/mmol)", .) %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
     gsub('Age', 'Age (years)', .) %>%

### 2017-11-06 doc/vitaminD.Rmd
formatted graph outputs to match with slides colour theme

@@
 vitd %>% 
-  box_plot("vitdStatus", "log(udbpCr)", 
+  box_plot_slides("vitdStatus", "log(udbpCr)", 
             "Vitamin D Status",
@@
 dsBase %>% 
-  scatter_plot("log(udbpCr)", "VitaminD", 
+  scatter_plot_slides("log(udbpCr)", "VitaminD", 
                "log UDBP:Creatinine",

### 2017-11-06 R/plots.R

@@ -81,17 +81,16 @@ box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
                          alpha = 0.3,
                          size = 1,
                          colour = "grey30") +
-    ggplot2::theme_minimal() +
-    ggplot2::scale_color_manual(values = c("#0db7c4", "#f24745", "#a9d039")) +
-    ggplot2::scale_fill_manual(values = c("#0db7c4", "#f24745", "#a9d039")) +
+    ggplot2::scale_color_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
+    ggplot2::scale_fill_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
     ggplot2::theme(legend.position = "none",
+                   panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 14, family = "Open Sans"),
-                   axis.text.y = ggplot2::element_text(colour = "grey30"),
+                   text = ggplot2::element_text(size = 14, colour = "#2b3443"),
+                   axis.text.y = ggplot2::element_text(colour = "#2b3443"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "grey30"),
-                   axis.title = ggplot2::element_text(family = "Dosis")) + #angle = 45
+                   axis.text.x = ggplot2::element_text(colour = "#2b3443")) +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
@@ -136,7 +135,7 @@ box_plot_poster <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
                    plot.background = ggplot2::element_rect(colour = "grey60"),
-                   text = ggplot2::element_text(size = 14, family = "Calibri Light"),
+                   # text = ggplot2::element_text(size = 14, family = "Open Sans"),
                    axis.text.y = ggplot2::element_text(colour = "grey30"),
[…]

### 2017-11-08 doc/vitaminD.Rmd
explorary analysis with eGFR and uVDBP versus uVDBP:creatinine

@@
   filter(fVN == "Baseline") %>% 
-  select(vitdStatus, udbpCr) %>%
+  select(vitdStatus, udbpCr, UDBP) %>%
   na.omit()
@@
 vitd %>% 
-  box_plot_slides("vitdStatus", "log(udbpCr)", 
+  box_plot("vitdStatus", "log(UDBP)", 
             "Vitamin D Status",
-            "log uVDBP:Creatinine (μg/mmol)")
+            "log uVDBP (μg/mmol)")
 

### 2017-11-08 R/fetch_data.R

@@ -32,6 +32,7 @@ fetch_data <- function() {
         labels = c("Baseline", "3Year", "6Year"),
         ordered = TRUE
       ),
+      YearsFromBaseline = MonthsFromBaseline/12,
       fMedsBP = factor(
         MedsBloodPressure,
         levels = c(0, 1),
@@ -95,6 +96,7 @@ fetch_data <- function() {
       VN,
       fVN,
       MonthsFromBaseline,
+      YearsFromBaseline,
       VisitDate,
       BMI,
       Waist,

### 2017-11-08 R/gee.R

@@ -15,9 +15,6 @@ prep_mason_data <- function(data) {
       UDBP = UDBP/1000, # udbp units to ug/mL
       udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
       ageBase = ifelse(fVN == "Baseline", Age, NA),
-      DM = ifelse(DM == 1, "DM", "notDM"),
-      fDM = relevel(as.factor(DM), "notDM"),
-      fDysglycemia = ifelse(!(dmStatus == "NGT"), "Dysglycemia", "notDysglycemia"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other"),
       dmStatus = factor(dmStatus, ordered = FALSE)
@@ -25,7 +22,6 @@ prep_mason_data <- function(data) {
     dplyr::arrange(SID, fVN) %>%
     dplyr::group_by(SID) %>%
     tidyr::fill(udbpCrBase, ageBase) %>%
-    # dplyr::mutate_at(dplyr::vars(-SID, -fVN), dplyr::funs(as.numeric(scale(.)))) %>%
     dplyr::ungroup() %>%
     dplyr::arrange(SID, VN)
 }
@@ -42,8 +38,9 @@ prep_mason_data_kidney <- function(data) {
       udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
       udbpCrBase = ifelse(fVN == "Baseline", udbpCr, NA),
       ageBase = ifelse(fVN == "Baseline", Age, NA),
-      DM = ifelse(DM == 1, "DM", "notDM"),
-      fDM = relevel(as.factor(DM), "notDM"),
+      acrBase = ifelse(fVN == "Baseline", ACR, NA),
+      scUDBP = UDBP/1000, # udbp units to ug/mL,
+      scudbpCr = scUDBP/UrineCreatinine,
       dmStatus = factor(dmStatus, ordered = FALSE),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other"),
@@ -59,15 +56,15 @@ prep_mason_data_kidney <- function(data) {
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
                        UDBP,
-                       udbpBase,
                        udbpCr,
+                       udbpBase,
                        udbpCrBase,
-                       MonthsFromBaseline,
+                       acrBase,
[…]

### 2017-11-08 R/plots.R

@@ -34,11 +34,11 @@ box_plot <- function(data=project_data, xvar, yvar, xlab="", ylab="", facet = FA
                          alpha = 0.3,
                          size = 1,
                          colour = "grey30") +
-    ggplot2::theme_minimal() +
-    ggplot2::scale_fill_brewer(palette = "Dark2") +
-    ggplot2::scale_color_brewer(palette = "Dark2") +
+    ggplot2::scale_color_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
+    ggplot2::scale_fill_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
     ggplot2::theme(legend.position = "none",
                    panel.grid.major.x = ggplot2::element_blank(),
+                   panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
                    axis.line.y = ggplot2::element_blank(),
                    axis.text.y = ggplot2::element_text(colour = "grey"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
@@ -168,30 +168,25 @@ box_plot_poster <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
 #'
 #' @examples
 scatter_plot = function(data=project_data,
-                        xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
-
+                               xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
   myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
-    ggplot2::geom_point(size = 1.5, alpha = 0.8) + #mapping=aes(color=mcr_status), #0db7c4
-    ggplot2::scale_fill_brewer(palette = "Dark2") +
-    ggplot2::scale_color_brewer(palette = "Dark2") +
-    ggplot2::theme_minimal() +
+    ggplot2::geom_point(colour = "#2b3443", size = 1) +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    panel.grid.minor.x = ggplot2::element_blank(),
+                   panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 11), #family = "Open Sans"
[…]

### 2017-11-08 R/rename.R

@@ -48,6 +48,7 @@ rename_gee_kidney <- function(x) {
   x %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
+    gsub("YearsFromBaseline", "Follow-up Duration (years)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
     gsub('leGFR', 'eGFR (ml/min/1.73m^2)', .) %>%

### 2017-11-08 doc/vitaminD.Rmd
vitamin D results for committee meeting document

@@
 vitd %>% 
-  box_plot("vitdStatus", "log(UDBP)", 
+  box_plot("vitdStatus", "log(udbpCr)", 
             "Vitamin D Status",
-            "log uVDBP (μg/mmol)")
+            "log uVDBP:cr (μg/mmol)")
 
@@
 dsBase %>% 
-  scatter_plot_slides("log(udbpCr)", "VitaminD", 
+  scatter_plot("log(udbpCr)", "VitaminD", 
                "log UDBP:Creatinine",
@@
 
-plot_gee_results_vitd(gee_vitd_baseline,
+plot_gee_results_vitd_base(gee_vitd_baseline,
                  yvars = "Serum 25(OH)D (nmol/L)")
@@
 plot_gee_results_vitd(gee_vitd,
-                 yvars = c("Serum 25(OH)D (nmol/L)",
-                           "PTH (pmol/L)"))
+                 yvars = "Serum 25(OH)D (nmol/L)")
 

### 2017-11-09 doc/vitaminD.Rmd
cleaned up gee output using functions

@@
 dsBase %>% 
-  scatter_plot("log(udbpCr)", "VitaminD", 
-               "log UDBP:Creatinine",
-               "Serum 25(OH)D")
+  scatter_plot("log(VitaminD)", "log(udbpCr)", 
+               "log serum 25(OH)D",
+               "log UDBP:Creatinine")
 
@@
 dsBase %>% 
-  cor.test(formula = ~ VitaminD + udbpCr, data = ., method = "spearman")
+  cor.test(formula = ~ log(udbpCr) + log(VitaminD), data = ., method = "spearman")
 
@@
 dsBase %>% 
+  dplyr::mutate(lVitD = log(VitaminD)) %>% 
   prep_mason_data() %>% 
-  mason_glm(y = "VitaminD",
-            x = "logudbpCr",
-            covars = c("ageBase", "Sex", "Ethnicity", "BMI", "MET", "fDM")
+  mason_glm(y = "logudbpCr",
+            x = "lVitD",
+            covars = c("ageBase", "Sex", "Ethnicity", "BMI", "MET", "Season", "dmStatus")
             ) %>% 
-  dplyr::filter(!term == "(Intercept)") %>%
-  dplyr::mutate(p = round(p.value, 2),
-    p = ifelse(p == "0", "<0.001", p),
-    estCI = paste0(round(estimate, 2), " (",
-                               round(conf.low, 2), ", ",
-                               round(conf.high, 2), ")")) %>% 
-  dplyr::select(Yterms, Xterms, term, estCI, p) %>% 
-  # tidyr::spread(Yterms, estCI) %>%
+  gee_results_table() %>% 
   pander::pander()

### 2017-11-09 R/gee.R
exploratory analysis for uVDBP vs uVDBP:cr

@@ -255,7 +255,7 @@ plot_gee_results_kidney_base <- function(results, yvars,
                                "EthnicityEuropean",
                                "dmStatusPreDiabetes",
                                "dmStatusDiabetes"),
-                     labels = c("Baseline ACR",
+                     labels = c("Baseline UDBP:cr",
                                 "Follow-up Duration (Years)",
                                 "Baseline Age (Years)",
                                 "Sex (male)",

### 2017-11-16 R/fetch_data.R
removed 15 subjects taking oral contraceptives from analysis

@@ -94,6 +94,7 @@ fetch_data <- function() {
     dplyr::filter(eGFR < 200) %>%
     dplyr::filter(eGFR_mdrd < 300) %>%
     dplyr::filter(Creatinine < 200) %>%
+    dplyr::filter(OralContraceptive == 1) %>%
     dplyr::select(
       SID,
       VN,

### 2017-11-17 R/fetch_data.R
remove oral contraceptive subjects

@@ -94,7 +94,8 @@ fetch_data <- function() {
     dplyr::filter(eGFR < 200) %>%
     dplyr::filter(eGFR_mdrd < 300) %>%
     dplyr::filter(Creatinine < 200) %>%
-    dplyr::filter(OralContraceptive == 1) %>%
+    dplyr::filter(OralContraceptive == 0) | (is.na(OralContraceptive)) %>%
+    # dplyr::filter(is.na(OralContraceptive)) %>%
     dplyr::select(
       SID,
       VN,
@@ -158,7 +159,7 @@ fetch_data <- function() {
       # dplyr::matches("meds")
     )
 
-  # Final dataset object
+   # Final dataset object
   project_data <- ds.prep
 
   # Save the dataset to the data/ folder.

### 2017-11-22 R/plots.R
aesthetic improvements to sample size label

@@ -71,7 +71,7 @@ box_plot <- function(data=project_data, xvar, yvar, xlab="", ylab="", facet = FA
 box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", facet = FALSE) {
 
   give.n <- function(x){
-    return(c(y = mean(x), label = length(x)))
+    return(data.frame(y = max(x) + 1, label = paste0("n = ",length(x))))
   }
 
   myboxplot <- ggplot2::ggplot(data, ggplot2::aes_string(x = xvar, y = yvar)) +
@@ -81,7 +81,7 @@ box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
                           fun.data = function(x){
                             return(c(y = median(x), ymin = median(x), ymax = median(x)))
                           }) +
-    ggplot2::stat_summary(fun.data = give.n, geom = "text") +
+    ggplot2::stat_summary(fun.data = give.n, geom = "text", size = 4, colour = "grey") +
     ggplot2::geom_jitter(ggplot2::aes_string(),
                          position = ggplot2::position_jitter(width = 0.15, height = 0),
                          alpha = 0.3,

### 2017-11-27 R/fetch_data.R
keep subjects with oral contraceptives

@@ -94,7 +94,7 @@ fetch_data <- function() {
     dplyr::filter(eGFR < 200) %>%
     dplyr::filter(eGFR_mdrd < 300) %>%
     dplyr::filter(Creatinine < 200) %>%
-    dplyr::filter(!(fVN == "Baseline" & OralContraceptive == 1)) %>%
+    # dplyr::filter(!(fVN == "Baseline" & OralContraceptive == 1)) %>%
     # dplyr::filter(is.na(OralContraceptive)) %>%
     dplyr::select(
       SID,

### 2017-11-27 R/plots.R
configured plots seminar slides

@@ -84,8 +84,8 @@ box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
     ggplot2::stat_summary(fun.data = give.n, geom = "text", size = 4, colour = "grey") +
     ggplot2::geom_jitter(ggplot2::aes_string(),
                          position = ggplot2::position_jitter(width = 0.15, height = 0),
-                         alpha = 0.3,
-                         size = 1,
+                         alpha = 0.1,
+                         size = 2,
                          colour = "grey30") +
     ggplot2::scale_color_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
     ggplot2::scale_fill_manual(values = c("#2b3443", "#adc9b7", "#ffd91e")) +
@@ -218,8 +218,9 @@ scatter_plot = function(data=project_data,
 #' @examples
 scatter_plot_slides = function(data=project_data,
                                xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
+
   myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
-    ggplot2::geom_point(colour = "#2b3443", size = 1) +
+    ggplot2::geom_point(colour = "#2b3443", shape = 16, size = 3, alpha = 0.15, show.legend = FALSE) +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    panel.grid.minor.x = ggplot2::element_blank(),
                    panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
@@ -228,14 +229,15 @@ scatter_plot_slides = function(data=project_data,
                    axis.text.y = ggplot2::element_text(colour = "#2b3443"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
                    axis.text.x = ggplot2::element_text(colour = "#2b3443")) +
+    # ggplot2::scale_color_gradient(low = "#0091ff", high = "#f0650e") +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
   if(line == TRUE & facet == FALSE) {
     myplot +
-      ggplot2::geom_smooth(method = lm, colour = "#ffd91e", fill = "#ffd91e")
+      ggplot2::geom_smooth(method = lm, colour = "#ffd91e", fill = "#ffd91e", alpha = 0.25)
   } else if(line == TRUE & facet == TRUE) {
     myplot +
[…]

### 2017-11-27 R/gee.R
removed space in x label for plots

@@ -320,7 +320,7 @@ plot_gee_results_kidney <- function(results, yvars,
 #'
 #' @examples
 plot_gee_results_vitd_base <- function(results, yvars,
-                                    xlab = "Percent difference with 95% CI in the outcomes for\n
+                                    xlab = "Percent difference with 95% CI in the outcomes for
                                   each SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
@@ -335,7 +335,6 @@ plot_gee_results_vitd_base <- function(results, yvars,
                                                  "SexMale",
                                                  "EthnicityEuropean",
                                                  "MET (kcal/kg/h)",
-                                                 "BMI (kg/m^2)",
                                                  "dmStatusPreDiabetes",
                                                  "dmStatusDiabetes",
                                                  "SeasonWinter",
@@ -346,7 +345,6 @@ plot_gee_results_vitd_base <- function(results, yvars,
                                                  "Sex (male)",
                                                  "Ethnicity (European)",
                                                  "MET (kcal/kg/h)",
-                                                 "BMI (kg/m^2)",
                                                  "Prediabetes",
                                                  "Diabetes",
                                                  "Seasonality (winter)",
@@ -358,7 +356,7 @@ plot_gee_results_vitd_base <- function(results, yvars,
 
 
 plot_gee_results_vitd <- function(results, yvars,
-                                  xlab = "Percent difference with 95% CI in the outcomes for\n
+                                  xlab = "Percent difference with 95% CI in the outcomes for
                                   each SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
@@ -373,7 +371,6 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "SexMale",
                                                  "EthnicityEuropean",
                                                  "MET (kcal/kg/h)",
-                                                 "BMI (kg/m^2)",
                                                  "dmStatusPreDiabetes",
                                                  "dmStatusDiabetes",
                                                  "SeasonWinter",
@@ -384,7 +381,6 @@ plot_gee_results_vitd <- function(results, yvars,
                                                  "Sex (male)",
                                                  "Ethnicity (European)",
                                                  "MET (kcal/kg/h)",
-                                                 "BMI (kg/m^2)",
                                                  "Prediabetes",
                                                  "Diabetes",
                                                  "Seasonality (winter)",

### 2017-11-27 doc/vitaminD.Rmd
adjusted figures to match seminar slides

@@
 dsBase %>% 
-  scatter_plot("log(VitaminD)", "log(udbpCr)", 
+  scatter_plot_slides("log(VitaminD)", "log(udbpCr)", 
                "log serum 25(OH)D",
-               "log UDBP:Creatinine")
+               "log uVDBP:cr (μg/mmol)")
 
@@
             x = "lVitD",
-            covars = c("ageBase", "Sex", "Ethnicity", "BMI", "MET", "Season", "dmStatus")
+            covars = c("ageBase", "Sex", "Ethnicity", "MET", "Season", "dmStatus")
             ) %>% 
@@
             covars = c("MonthsFromBaseline", "ageBase", "Sex", 
-                       "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+                       "Ethnicity", "MET", "dmStatus", "Season"),
             intvar = "Season") %>% 
@@
             x = "udbpCr",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "BMI", "dmStatus", "Season"),
+            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "dmStatus", "Season"),
             intvar = "Season") %>% 

### 2017-12-04 R/fetch_data.R
missingness for glycemic status

@@ -108,8 +108,8 @@ fetch_data <- function() {
       Age,
       Sex,
       Ethnicity,
-      Glucose0,
-      Glucose120,
+      Glucose0_OGTT,
+      Glucose120_OGTT,
       DM,
       IFG,
       IGT,

### 2017-12-04 R/fetch_data.R
added dysgly variable to analysis

@@ -110,6 +110,7 @@ fetch_data <- function() {
       Ethnicity,
       Glucose0_OGTT,
       Glucose120_OGTT,
+      Dysgly,
       DM,
       IFG,
       IGT,

### 2017-12-06 R/gee.R
made plots more visible for seminar presentation

@@ -246,8 +246,7 @@ gee_results_table <- function(results, table = TRUE) {
 #'
 #' @examples
 plot_gee_results_kidney_base <- function(results, yvars,
-                     xlab = "Percent difference with 95% CI in the outcomes
-                     for each SD increase in ACR and covariates",
+                     xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in ACR and covariates",
                      terms = c("<-Xterm",
                                "Follow-up Duration (years)",
                                "Baseline Age (years)",
@@ -255,7 +254,7 @@ plot_gee_results_kidney_base <- function(results, yvars,
                                "EthnicityEuropean",
                                "dmStatusPreDiabetes",
                                "dmStatusDiabetes"),
-                     labels = c("Baseline UDBP:cr",
+                     labels = c("Baseline uVDBP:cr",
                                 "Follow-up Duration (Years)",
                                 "Baseline Age (Years)",
                                 "Sex (male)",
@@ -278,8 +277,7 @@ plot_gee_results_kidney_base <- function(results, yvars,
 
 
 plot_gee_results_kidney <- function(results, yvars,
-                                         xlab = "Percent difference with 95% CI in the outcomes
-                                         for each SD increase in uVDBP and covariates",
+                                         xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates",
                                          terms = c("<-Xterm",
                                                    "Follow-up Duration (years)",
                                                    "Baseline Age (years)",
@@ -320,8 +318,7 @@ plot_gee_results_kidney <- function(results, yvars,
 #'
 #' @examples
 plot_gee_results_vitd_base <- function(results, yvars,
-                                    xlab = "Percent difference with 95% CI in the outcomes for
-                                  each SD increase in uVDBP and covariates") {
+                                    xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
     dplyr::filter(!term == "(Intercept)") %>%
@@ -356,8 +353,7 @@ plot_gee_results_vitd_base <- function(results, yvars,
 
 
 plot_gee_results_vitd <- function(results, yvars,
-                                  xlab = "Percent difference with 95% CI in the outcomes for
-                                  each SD increase in uVDBP and covariates") {
[…]

### 2017-12-06 R/plots.R

@@ -93,10 +93,10 @@ box_plot_slides <- function(data=project_data, xvar, yvar, xlab="", ylab="", fac
                    panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
                    panel.grid.major.x = ggplot2::element_blank(),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 14, colour = "#2b3443"),
-                   axis.text.y = ggplot2::element_text(colour = "#2b3443"),
-                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "#2b3443")) +
+                   text = ggplot2::element_text(size = 14, colour = "grey30"),
+                   axis.text.y = ggplot2::element_text(colour = "grey50"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey70"),
+                   axis.text.x = ggplot2::element_text(colour = "grey50")) +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
 
@@ -225,10 +225,10 @@ scatter_plot_slides = function(data=project_data,
                    panel.grid.minor.x = ggplot2::element_blank(),
                    panel.background = ggplot2::element_rect(fill = "#f5f3ef", colour = "#f5f3ef"),
                    axis.line.y = ggplot2::element_blank(),
-                   text = ggplot2::element_text(size = 14, colour = "#2b3443"),
-                   axis.text.y = ggplot2::element_text(colour = "#2b3443"),
-                   axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "#2b3443")) +
+                   text = ggplot2::element_text(size = 14, colour = "grey30"),
+                   axis.text.y = ggplot2::element_text(colour = "grey50"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey70"),
+                   axis.text.x = ggplot2::element_text(colour = "grey50")) +
     # ggplot2::scale_color_gradient(low = "#0091ff", high = "#f0650e") +
     ggplot2::xlab(xlab) +
     ggplot2::ylab(ylab)
@@ -383,10 +383,10 @@ gee_plot = function(gee_results, xlab = "") {
                    strip.text.x = ggplot2::element_text(colour = "white", face = "bold", size = 11),
[…]

### 2018-03-01 R/fetch_data.R
merged data sets with meds, bloodwork, ogtt, and misc into PROMISE

@@ -14,6 +14,9 @@ fetch_data <- function() {
   # Load the master dataset,
   ds.prep <- PROMISE.data::PROMISE %>%
 
+  # Merge data sets
+    PROMISE.data::combine_datasets(meds, bloodwork, ogtt, misc) %>%
+
   # Data wrangling commands
     dplyr::mutate(
       UDBP = ifelse(UDBP == 0.01, NA, UDBP),
@@ -38,9 +41,8 @@ fetch_data <- function() {
         levels = c(0, 1),
         labels = c("No", "Yes")
       ),
-      Hypertension = ifelse(!(is.na(MedHistHypertenAge)), 1,
-                            ifelse(Systolic > 140 | Diastolic > 90, 1,
-                                   0)),
+      Hypertension = ifelse(Systolic > 140 | Diastolic > 90, 1,
+                                   0),
       Hypertension = factor(Hypertension,
                             levels = c(0, 1),
                             labels = c("No", "Yes")),
@@ -86,7 +88,6 @@ fetch_data <- function() {
                        breaks = c(-Inf, 50, 75, Inf),
                        labels = c("Deficient", "Insufficient", "Sufficient"),
                        ordered = TRUE),
-      OralContraceptive = ifelse(BirthControl == "2", 1, 0),
       Season = ifelse(lubridate::month(VisitDate) %in% c("5", "6", "7", "8", "9", "10"),
                       "Summer", "Winter")
     ) %>%
@@ -145,11 +146,6 @@ fetch_data <- function() {
       SmokeCigs,
       CRP,
       Canoe,
-      BirthControl,
-      OralContraceptive,
-      PeriodsStopped,
-      PeriodsStoppedAge,
-      MedHistHypertenAge,
       MedsBloodPressure,
       MedsDxDM,
       diet_milk,

### 2018-03-01 R/plots.R
adjusted GEE plot to a smaller font

@@ -381,12 +381,12 @@ gee_plot = function(gee_results, xlab = "") {
                    panel.spacing = ggplot2::unit(2, "lines"),
                    strip.background = ggplot2::element_rect(colour = "#adc9b7", fill = "#adc9b7"),
                    strip.text.x = ggplot2::element_text(colour = "white", face = "bold", size = 11),
-                   # text = ggplot2::element_text(size = 14, colour = "#2b3443"),
+                   text = ggplot2::element_text(size = 11, colour = "#2b3443"),
                    axis.line.y = ggplot2::element_blank(),
-                   axis.text.y = ggplot2::element_text(size = 12, colour = "grey50"),
+                   axis.text.y = ggplot2::element_text(size = 10, colour = "grey50"),
                    axis.ticks.y = ggplot2::element_line(colour = "grey"),
-                   axis.text.x = ggplot2::element_text(colour = "grey40"),
-                   axis.title = ggplot2::element_text(size = 14, colour = "#2b3443")) + #family = "Montserrat"
+                   axis.text.x = ggplot2::element_text(size = 11, colour = "grey40"),
+                   axis.title = ggplot2::element_text(size = 11, colour = "#2b3443")) + #family = "Montserrat"
     ggplot2::ylab('') +
     ggplot2::xlab(xlab)
 }

### 2018-03-12 R/gee.R
x axis label changed to be on one line

@@ -277,7 +277,7 @@ plot_gee_results_kidney_base <- function(results, yvars,
 
 
 plot_gee_results_kidney <- function(results, yvars,
-                                         xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates",
+                                         xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates",
                                          terms = c("<-Xterm",
                                                    "Follow-up Duration (years)",
                                                    "Baseline Age (years)",

### 2018-03-12 R/plots.R
removed border around plot area

@@ -267,7 +267,7 @@ scatter_plot_poster = function(data=project_data,
     ggplot2::geom_point(colour = "#adc9b7", size = 1) +
     ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
                    panel.grid.minor.x = ggplot2::element_blank(),
-                   panel.background = ggplot2::element_rect(colour = "grey60"),
+                   # panel.background = ggplot2::element_rect(colour = "grey60"),
                    axis.line.y = ggplot2::element_blank(),
                    text = ggplot2::element_text(size = 14, family = "Montserrat Light"),
                    axis.text.y = ggplot2::element_text(colour = "grey30"),
@@ -278,10 +278,10 @@ scatter_plot_poster = function(data=project_data,
     ggplot2::ylab(ylab)
   if(line == TRUE & facet == FALSE) {
     myplot +
-      ggplot2::geom_smooth(method = loess, colour = "grey50")
+      ggplot2::geom_smooth(method = lm, colour = "grey50")
   } else if(line == TRUE & facet == TRUE) {
     myplot +
-      ggplot2::geom_smooth(method = loess, colour = "grey50") +
+      ggplot2::geom_smooth(method = lm, colour = "grey50") +
       ggplot2::facet_grid(~fVN, switch = "x")
   } else if(line == FALSE & facet == TRUE) {
     myplot +

### 2018-03-12 doc/adaAbstract.Rmd
updated variable names

@@
 knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
-library(dplyr)
+library(tidyverse)
 devtools::load_all()
@@
   na.omit() %>%
-  box_plot_poster("acrStatus", "log(udbpCrRatio)", 
+  box_plot_poster("acrStatus", "log(udbpCr)", 
             "Albuminuria",
@@
   filter(fVN == "Baseline") %>% 
-  scatter_plot_poster("log(ACR)", "log(udbpCrRatio)",
+  scatter_plot_poster("log(ACR)", "log(udbpCr)",
                "log Albumin:Creatinine Ratio",
@@
   filter(fVN == "Baseline") %>% 
-  cor.test(formula = ~ ACR + udbpCrRatio, data = ., method = "spearman")
+  cor.test(formula = ~ ACR + udbpCr, data = ., method = "spearman")
 ```
@@
   dplyr::filter(fVN == "Baseline") %>% 
-  scatter_plot_poster("eGFR", "log(udbpCrRatio)",
+  scatter_plot_poster("eGFR", "log(udbpCr)",
                "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
@@
 ds %>% 
-  filter(!(acr_status == "Macroalbuminuria")) %>% 
-  cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")
+  filter(!(acrStatus == "Macroalbuminuria")) %>% 
+  cor.test(formula = ~ eGFR + udbpCr, data = ., method = "spearman")
 ```

### 2018-03-16 R/gee.R
do not remove dm from baseline in gee models

@@ -52,7 +52,7 @@ prep_mason_data_kidney <- function(data) {
     dplyr::filter(!(fVN == "Baseline" &
                       acrStatus == "Macroalbuminuria")) %>%
     dplyr::filter(!(fVN == "Baseline" & eGFR < 60)) %>%
-    dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
+    # dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
                        UDBP,
@@ -96,8 +96,8 @@ prep_mason_data_vitd <- function(data) {
       # lPTH = log(PTH)
     ) %>%
 
-    # dplyr::filter(!(fVN == "Baseline" & vitdStatus == "Deficient")) %>%
-    dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
+    dplyr::filter(!(fVN == "Baseline" & vitdStatus == "Deficient")) %>%
+    # dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
                        UDBP,
@@ -318,7 +318,7 @@ plot_gee_results_kidney <- function(results, yvars,
 #'
 #' @examples
 plot_gee_results_vitd_base <- function(results, yvars,
-                                    xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates") {
+                                    xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
     dplyr::filter(!term == "(Intercept)") %>%
@@ -353,7 +353,7 @@ plot_gee_results_vitd_base <- function(results, yvars,
 
 
 plot_gee_results_vitd <- function(results, yvars,
-                                  xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates") {
+                                  xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
     dplyr::filter(!term == "(Intercept)") %>%

### 2018-03-16 R/gee.R
added BMI covariate in vitamin D GEE model

@@ -90,6 +90,7 @@ prep_mason_data_vitd <- function(data) {
       fDM = relevel(as.factor(DM), "notDM"),
       Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
       Ethnicity = relevel(as.factor(Ethnicity), "Other"),
+      Season = relevel(as.factor(Season), "Winter"),
       dmStatus = factor(dmStatus, ordered = FALSE),
 
       lVitD = log(VitaminD)
@@ -327,25 +328,27 @@ plot_gee_results_vitd_base <- function(results, yvars,
                                   ordered = TRUE),
                   Xterms = factor(Xterms,
                                   levels = rev(c("<-Xterm",
-                                                 "Follow-up Duration (months)",
+                                                 "Follow-up Duration (years)",
                                                  "Baseline Age (years)",
                                                  "SexMale",
                                                  "EthnicityEuropean",
+                                                 "BMI (kg/m^2)",
                                                  "MET (kcal/kg/h)",
                                                  "dmStatusPreDiabetes",
                                                  "dmStatusDiabetes",
-                                                 "SeasonWinter",
-                                                 "<-Xterm:SeasonWinter")),
+                                                 "SeasonSummer")),
+                                                 # "<-Xterm:SeasonWinter")),
                                   labels = rev(c("Baseline uVDBP:cr (ug/mL)",
-                                                 "Follow-up Duration (months)",
+                                                 "Follow-up Duration (years)",
                                                  "Baseline Age (years)",
                                                  "Sex (male)",
                                                  "Ethnicity (European)",
+                                                 "BMI (kg/m^2)",
                                                  "MET (kcal/kg/h)",
                                                  "Prediabetes",
                                                  "Diabetes",
-                                                 "Seasonality (winter)",
-                                                 "Season:uVDBP:cr interaction")),
+                                                 "Seasonality (summer)")),
+                                                 # "Season:uVDBP:cr interaction")),
                                   ordered = TRUE)) %>%
     arrange(Xterms) %>%
     gee_plot(xlab = xlab)
@@ -362,25 +365,27 @@ plot_gee_results_vitd <- function(results, yvars,
                                   ordered = TRUE),
                   Xterms = factor(Xterms,
                                   levels = rev(c("<-Xterm",
-                                                 "Follow-up Duration (months)",
+                                                 "Follow-up Duration (years)",
                                                  "Baseline Age (years)",
                                                  "SexMale",
[…]

### 2018-03-16 R/rename.R
added yearsfrombaseline

@@ -8,8 +8,9 @@
 #' @examples
 rename_variables <- function(x) {
   x %>%
-    gsub("VN", "Follow-up Duration (years)", .) %>%
+    gsub("VN", "Visit Number", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
+    gsub("YearsFromBaseline", "Follow-up Duration (years)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
     gsub("BMI", "BMI (kg/m^2)", .) %>%
@@ -76,6 +77,7 @@ rename_gee_vitd <- function(x) {
     gsub("VitaminD", "Serum 25(OH)D (nmol/L)", .) %>%
     gsub("VN", "Follow-up Duration (years)", .) %>%
     gsub("MonthsFromBaseline", "Follow-up Duration (months)", .) %>%
+    gsub("YearsFromBaseline", "Follow-up Duration (years)", .) %>%
     gsub('Age', 'Age (years)', .) %>%
     gsub("ageBase", "Baseline Age (years)", .) %>%
     gsub("BMI", "BMI (kg/m^2)", .) %>%

### 2018-03-16 doc/vitaminD.Rmd
made plots more visible for seminar presentation
modified GEE model to remove season and include BMI

@@
 devtools::load_all()
-load_data(update = TRUE)
+# load_data(update = TRUE)
 set_options()
@@
   dplyr::group_by(vitdStatus) %>% 
-  dplyr::summarise(Median = median(log(udbpCr)), 
+  dplyr::summarise(n = n(),
+                   Median = median(log(udbpCr)), 
+                   lower = quantile(log(udbpCr), probs = 0.25),
+                   higher = quantile(log(udbpCr), probs = 0.75),
                    IQR = IQR(log(udbpCr))
@@
   scatter_plot_slides("log(VitaminD)", "log(udbpCr)", 
-               "log serum 25(OH)D",
+               "log serum 25(OH)D (nmol/L)",
                "log uVDBP:cr (μg/mmol)")
@@
             x = "udbpCrBase",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", 
-                       "Ethnicity", "MET", "dmStatus", "Season"),
-            intvar = "Season") %>% 
+            covars = c("YearsFromBaseline", "ageBase", "Sex", 
+                       "Ethnicity", "BMI", "MET", "dmStatus", "Season")) %>% 
+            # intvar = "Season") %>% 
   mason::polish_renaming(rename_gee_vitd)
@@
 # GEE table ----------------------------------------------------------
-
-tab_nums(gee_vitd_baseline, "GEE results for uVDBP at baseline (predictor) with 25(OH)D and (outcomes), adjusted for follow-up duration (in months), baseline age, sex, ethnicity, physical activity (MET), BMI, diabetes status, and seasonality; subjects with deficient vitamin D status at baseline excluded (n = 713).")
 
@@
             x = "udbpCr",
-            covars = c("MonthsFromBaseline", "ageBase", "Sex", "Ethnicity", "MET", "dmStatus", "Season"),
-            intvar = "Season") %>% 
+            covars = c("YearsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI",
+                       "MET", "dmStatus", "Season")) %>% 
+            # intvar = "Season") %>% 
   mason::polish_renaming(rename_gee_vitd)

### 2018-03-27 R/gee.R
subgroup analysis with only kidney dysfunction subjects

@@ -98,6 +98,8 @@ prep_mason_data_vitd <- function(data) {
     ) %>%
 
     dplyr::filter(!(fVN == "Baseline" & vitdStatus == "Deficient")) %>%
+    # dplyr::filter(acrStatus != "Normoalbuminuria") %>%
+    # dplyr::filter(eGFRStatus != "Normal") %>%
     # dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),

### 2018-04-10 doc/vitaminD.Rmd
subgroup analysis with only kidney dysfunction subjects
determine n of subjects under assay detection limit

@@
 m <- ds %>% 
+  # dplyr::mutate(followUp = )
   dplyr::filter(fVN != "6Year")
@@
 dsBase %>% 
+  # dplyr::filter(acrStatus != "Normoalbuminuria") %>% 
   scatter_plot_slides("log(VitaminD)", "log(udbpCr)", 

### 2018-04-13 doc/vitaminD.Rmd
added kidney measures to table 5.2 (changes over time)

@@
            "PTH",
+           "ACR",
+           "eGFR",
            "UDBP",
@@
   strata = c("VN"),
-  data = m,
+  data = ds,
   factorVars = c("Sex", "Ethnicity", "dmStatus")
@@
   print(nonnormal = c("UDBP",
-                      "udbpCr"),
+                      "udbpCr",
+                      "ACR"),
         quote = FALSE,

### 2018-04-24 R/gee.R
removed scaling of follow-up duration in GEE

@@ -53,7 +53,7 @@ prep_mason_data_kidney <- function(data) {
     # dplyr::filter(!(fVN == "Baseline" & dmStatus == "DM")) %>%
 
     dplyr::mutate_each(dplyr::funs(as.numeric(scale(.))),
-                       YearsFromBaseline,
+                       # YearsFromBaseline,
                        UDBP,
                        udbpCr,
                        udbpBase,
@@ -248,7 +248,7 @@ gee_results_table <- function(results, table = TRUE) {
 #'
 #' @examples
 plot_gee_results_kidney_base <- function(results, yvars,
-                     xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in ACR and covariates",
+                     xlab = "Percent difference with 95% CI in the outcomes for each SD increase in ACR and covariates",
                      terms = c("<-Xterm",
                                "Follow-up Duration (years)",
                                "Baseline Age (years)",

### 2018-04-24 doc/vitaminD.Rmd
changed plot format to document style

@@
 vitd %>% 
-  box_plot_slides("vitdStatus", "log(udbpCr)", 
+  box_plot("vitdStatus", "log(udbpCr)", 
             "Vitamin D Status",
@@
   # dplyr::filter(acrStatus != "Normoalbuminuria") %>% 
-  scatter_plot_slides("log(VitaminD)", "log(udbpCr)", 
-               "log serum 25(OH)D (nmol/L)",
-               "log uVDBP:cr (μg/mmol)")
+  scatter_plot("log(udbpCr)", "log(VitaminD)", 
+                "log uVDBP:cr (μg/mmol)",
+               "log serum 25(OH)D (nmol/L)")
 

### 2018-05-03 R/gee.R
fixed line spacing on gee plots

@@ -320,7 +320,7 @@ plot_gee_results_kidney <- function(results, yvars,
 #'
 #' @examples
 plot_gee_results_vitd_base <- function(results, yvars,
-                                    xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates") {
+                                    xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
     dplyr::filter(!term == "(Intercept)") %>%
@@ -357,7 +357,7 @@ plot_gee_results_vitd_base <- function(results, yvars,
 
 
 plot_gee_results_vitd <- function(results, yvars,
-                                  xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates") {
+                                  xlab = "Percent difference with 95% CI in the outcomes for \neach SD increase in uVDBP and covariates") {
   results %>%
     dplyr::mutate(Xterms = term) %>%
     dplyr::filter(!term == "(Intercept)") %>%

### 2018-05-03 R/plots.R
scatter plot with transparent background

@@ -248,6 +248,39 @@ scatter_plot_slides = function(data=project_data,
 }
 
 
+
+scatter_plot_transparent = function(data=project_data,
+                               xvar, yvar, xlab='', ylab='', line = TRUE, facet = FALSE) {
+
+  myplot <- ggplot2::ggplot(data, ggplot2::aes_string(x=xvar, y=yvar)) +
+    ggplot2::geom_point(colour = "#2b3443", shape = 16, size = 3, alpha = 0.15, show.legend = FALSE) +
+    ggplot2::theme(panel.grid.major.x = ggplot2::element_blank(),
+                   panel.grid.minor.x = ggplot2::element_blank(),
+                   panel.background = ggplot2::element_rect(fill = "transparent", colour = NA),
+                   plot.background = ggplot2::element_rect(fill = "transparent", colour = NA),
+                   axis.line.y = ggplot2::element_blank(),
+                   text = ggplot2::element_text(size = 14, colour = "grey30"),
+                   axis.text.y = ggplot2::element_text(colour = "grey50"),
+                   axis.ticks.y = ggplot2::element_line(colour = "grey70"),
+                   axis.text.x = ggplot2::element_text(colour = "grey50")) +
+    # ggplot2::scale_color_gradient(low = "#0091ff", high = "#f0650e") +
+    ggplot2::xlab(xlab) +
+    ggplot2::ylab(ylab)
+  if(line == TRUE & facet == FALSE) {
+    myplot +
+      ggplot2::geom_smooth(method = loess, colour = "#ffd91e", fill = "#ffd91e", alpha = 0.25)
+  } else if(line == TRUE & facet == TRUE) {
+    myplot +
+      ggplot2::geom_smooth(method = lm, colour = "#ffd91e", fill = "#ffd91e", alpha = 0.25) +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else if(line == FALSE & facet == TRUE) {
+    myplot +
+      ggplot2::facet_grid(~fVN, switch = "x")
+  } else {
+    myplot
+  }
+}
+
 #' Scatterplot where the colour theme matches ada poster
 #'
 #' @param data Clearned data

### 2018-05-03 doc/vitaminD.Rmd
changed labels on boxplots to include values

@@
 # Clean data
-vitd <- ds %>% 
-  filter(fVN == "Baseline") %>% 
-  select(vitdStatus, udbpCr, UDBP) %>%
+vitd <- dsBase %>% 
+  dplyr::mutate(vitdLabels = ifelse(vitdStatus == "Sufficient", "Sufficient \n(≥75 nmol/L)",
+                                   ifelse(vitdStatus == "Insufficient", "Insufficient \n(50 - 75 nmol/L)",
+                                          "Deficient \n(<50 nmol/L)")),
+                vitdLabels = factor(vitdLabels,
+                                   levels = c("Sufficient \n(≥75 nmol/L)",
+                                              "Insufficient \n(50 - 75 nmol/L)",
+                                              "Deficient \n(<50 nmol/L)"),
+                                   ordered = TRUE)) %>%
+  select(vitdStatus, vitdLabels, udbpCr, UDBP) %>%
   na.omit()
@@
 vitd %>% 
-  box_plot("vitdStatus", "log(udbpCr)", 
+  box_plot_slides("vitdLabels", "log(udbpCr)", 
             "Vitamin D Status",
@@
   # dplyr::filter(acrStatus != "Normoalbuminuria") %>% 
-  scatter_plot("log(udbpCr)", "log(VitaminD)", 
+  scatter_plot_transparent("log(udbpCr)", "log(VitaminD)", 
                 "log uVDBP:cr (μg/mmol)",
@@
             covars = c("YearsFromBaseline", "ageBase", "Sex", 
-                       "Ethnicity", "BMI", "MET", "dmStatus", "Season")) %>% 
-            # intvar = "Season") %>% 
+                       "Ethnicity", "BMI", "MET", "dmStatus", "Season"),
+            intvar = "BMI") %>%
   mason::polish_renaming(rename_gee_vitd)
@@
             covars = c("YearsFromBaseline", "ageBase", "Sex", "Ethnicity", "BMI",
-                       "MET", "dmStatus", "Season")) %>% 
-            # intvar = "Season") %>% 
+                       "MET", "dmStatus", "Season"), 
+            intvar = "BMI") %>%
   mason::polish_renaming(rename_gee_vitd)

### 2018-05-15 doc/vitaminD.Rmd
changed vitamin D over time table to exclude those without vitamin D measures

@@
 m <- ds %>% 
-  # dplyr::mutate(followUp = )
-  dplyr::filter(fVN != "6Year")
+  dplyr::filter(!is.na(VitaminD)) %>% 
+  dplyr::filter(!VN == "6")
 
@@
   strata = c("VN"),
-  data = ds,
+  data = m,
   factorVars = c("Sex", "Ethnicity", "dmStatus")

### 2018-08-01 R/fetch_data.R
fixed data wrangling for creation of new variables (new raw data set variables updated)

@@ -15,9 +15,9 @@ fetch_data <- function() {
   ds.prep <- PROMISE.data::PROMISE %>%
 
   # Merge data sets
-    PROMISE.data::combine_datasets(meds, bloodwork, ogtt, misc) %>%
+    PROMISE.data::combine_datasets(ds.prep, meds, bloodwork, ogtt, misc) %>%
 
-  # Data wrangling commands
+  # Data wrangling
     dplyr::mutate(
       UDBP = ifelse(UDBP == 0.01, NA, UDBP),
       UrineCreatinine = ifelse(SID == 2028, 9, UrineCreatinine),
@@ -95,8 +95,6 @@ fetch_data <- function() {
     dplyr::filter(eGFR < 200) %>%
     dplyr::filter(eGFR_mdrd < 300) %>%
     dplyr::filter(Creatinine < 200) %>%
-    # dplyr::filter(!(fVN == "Baseline" & OralContraceptive == 1)) %>%
-    # dplyr::filter(is.na(OralContraceptive)) %>%
     dplyr::select(
       SID,
       VN,

### 2018-08-01 R/gee.R
adjustment to GEE model for eGFR (tested blood pressure medication)

@@ -308,6 +308,24 @@ plot_gee_results_kidney <- function(results, yvars,
     gee_plot(xlab = xlab)
 }
 
+plot_gee_results_eGFR <- function(results, yvars,
+                                    xlab = "Percent difference with 95% CI in the outcomes for each SD increase in uVDBP and covariates",
+                                    terms = covars,
+                                    labels = labels) {
+  results %>%
+    dplyr::mutate(Xterms = term) %>%
+    dplyr::filter(!term == "(Intercept)") %>%
+    dplyr::mutate(Yterms = factor(Yterms,
+                                  levels = yvars,
+                                  ordered = TRUE),
+                  Xterms = factor(Xterms,
+                                  levels = rev(terms),
+                                  labels = rev(labels),
+                                  ordered = TRUE)) %>%
+    arrange(Xterms) %>%
+    gee_plot(xlab = xlab)
+}
+
 
 #' Plot vitamin D GEE results in a forest plot-style
 #'
